---
format: html
execute:
  echo: false
  
bibliography: bibliography.bib
---

```{r}
#| label: set-up
#| echo: false
library(targets)
library(tarchetypes)
```

## Supporting Information

```{r}
#| label: tbl-sites
#| echo: false
#| tab-cap: Some text


tibble(Variable = c("Elevation m a.s.l.", "Latitude °N", "Longitude °E", "Mean summer temp. °C", "Annual precip in mm"),
         Lowland = c(469.0, 60.88019, 7.16990, 13, 1292),
         "Sub-alpine" = c(920.0, 60.86183, 7.16800, 10.4, 1256),
         Alpine = c(1290.0, 60.85994, 7.19504, 7.3, 2080)) |> 
  gt()
```

The alpine site was a species rich grassland dominated by *Silene acaulis*, *Thalictrum alpinum*, *Alchemilla alpina*, *Leontodon autumnalis*, *Antennaria dioica*, *Carex capillaris*, and *Festuca rubra*.
The sub-alpine site was located close to a summer farm and the vegetation was dominated by species adapted to grazing: *Achillea millefolium* , *Agrostis capillaris*, *Poa pratensis*, *Ranunculus repens*, and *Rumex acetosa* and the lowest site was located close to a farm and the vegetation is dominated by *Potentilla erecta*, *Knautia arvensis*, *Agrostis capillaris*, *Rumex acetosa*, *Festuca ovina*, and  *Poa pratense*.

