```{r}
#| label: setup-figures-main
#| echo: false
#| include: false

if (!dir.exists("_targets") && dir.exists("../_targets")) {
  targets::tar_config_set(store = normalizePath("../_targets", mustWork = TRUE))
}

library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(vegan)
library(patchwork)
```

![Two alternative hypotheses for how three global change drivers affect plant diversity. a) Warming, nitrogen deposition, and clipping (or grazing) affect diversity directly, or b) these drivers affect diversity indirectly, via their impacts on biotic interactions such as competition for light or nutrients, mediated through community biomass as a *common currency*.](figures/Figure1_hypotheses.png){#fig-concept width="600"}

![Experimental design and map of the study sites. We used a split-plot design to assess warming, nitrogen addition, clipping, and grazing, and replicated the study across two communities, an alpine and a sub-alpine plant community. Warmed communities were transplanted from the alpine to the sub-alpine and from the sub-alpine to the boreal site, in combination with nitrogen addition, clipping and grazing treatments. The treatments are ambient (grey), warming (red), nitrogen deposition (grey to green under ambient temperature and red to brown under warmed conditions, crossed with four levels of clipping and grazing (yellow); none (C), intermediate clipping (one scissor), intensive clipping (two scissors) and natural grazing (sheep). The arrows indicate the direction of transplant (warming treatment). Note that the nitrogen levels were applied at the plot level and the warming and clipping/grazing treatments were applied at the subplot level. Inset shows location of the study sites in south-western Norway.](figures/methods.jpg){#fig-methods width="800"}

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: "Community biomass (g m^-2^) (a) and Shannon diversity (b) with warming, clipping, and increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) community. The x axis is log transformed with numbers on the original scale for ease of interpretation. The colours indicate ambient (black) and warmed conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. Capital letters (black) within panels indicate significant model terms (W = warming, N = nitrogen addition, C = clipping, combined letters, e.g. WxC = interactions). Letters in grey indicate marginally significant model terms (0.05 < *P* < 0.07). R^2^ and Adj. R^2^ are shown in the upper right corner."

tar_read(bio_div_figure)
```

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping on Shannon diversity, as well as their indirect effects through community biomass for the a) alpine and b) sub-alpine communities. The numbers indicate the regression coefficients, the colour the direction (orange = positive, turquoise = negative),  and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05) and linetype  (solid = significant and dashed = non-significant)."

tar_load(cut_final_diversity)

cut_final_diversity[[1]]
```

```{r}
#| label: fig-toenails
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Distributions for species' environmental affinities (community weighted Ellenberg indicator values) at different levels of community biomass (a), and for warming (b) and nitrogen addition (c) treatments at the alpine (upper row) and sub-alpine communities (bottom row). Shown are the distributions for environmental affinities light, temperature, nutrient, moisture, salinity and reaction, and grazing disturbance. The colours indicate the mediator and different treatments: a) standing biomass, b) nitrogen addition levels, and c) warming (ambient = black, warming = red). Darker colours indicate higher values of biomass or nitrogen addition. The stars show the level of significance (* for P ≤ 0.05, ** for P ≤ 0.01, *** for P ≤ 0.001)."

tar_load(traits_biomass_plot)
tar_load(traits_warming_plot)
tar_load(traits_nitrogen_plot)

(traits_biomass_plot + labs(tag = "a) Community biomass")) +
    (traits_warming_plot + labs(tag = "b) Warming")) +
    (traits_nitrogen_plot + labs(tag = "c) Nitrogen")) +
    plot_layout(ncol = 1, guides = "collect") &
    theme(
        legend.position = "none",
        legend.box = "vertical",
        plot.tag.position = c(0.12, 1.08),
        plot.tag = element_text(size = 12, hjust = 0, vjust = 1),
        plot.margin = margin(t = 10, r = 5, b = 5, l = 5)
    )
```
