---
editor: visual
---

```{r}
#| label: set-up-methods
#| echo: false
#| message: false

library(targets)
library(tarchetypes)

if (!dir.exists("_targets") && dir.exists("../_targets")) {
  targets::tar_config_set(store = normalizePath("../_targets", mustWork = TRUE))
}

tar_load(standing_biomass_model_output)

```

## Material and Methods

### Study area and experimental design

Three semi-natural, species rich, perennial grassland sites (boreal, sub-alpine, and alpine) were established along an elevational gradient in south-western Norway (@fig-methods; 7.17°E, 60.88°N). The sites were approximately 400 m apart in elevation (469, 920, and 1290 m a.s.l.), corresponding to a temperature difference of approximately 3°C in summer air, ground, and soil temperature between the sites over the duration of the experiment (air: 3.09 ± 0.186 °C, ground: 3.06 ± 0.176 °C, and soil: 2.82 ± 0.142 °C; Figure S1, Table S1, S2). Annual precipitation was similar at the two lowest sites and approximately 800 mm higher at the alpine site. The sites were located on calcareous bedrock with loamy sand. Montane grasslands in Norway are traditionally managed as transhumance-based grazing systems [@reinton1955seter; @austad2004overview], and our sites were grazed by sheep, goats, and deer at the boreal and sub-alpine sites, and mainly by sheep and reindeer at the alpine site [@halbritter2025data]. Annual plant productivity is similar at the boreal and sub-alpine site and about four times lower at the alpine site, while species richness is approximately twice as high at the alpine site compared to the sub-alpine site. Grazing pressure is generally low, but 3 - 5 times higher in the sub-alpine site compared to the boreal and alpine sites [@halbritter2025effects]. Other factors such as land-use history, vegetation type, slope and exposure were kept as constant as possible within and across sites[@halbritter2025data]. To avoid accumulation of biomass, the biomass within the fenced experimental area was removed at the end of each field season.

In 2019, a global change experiment using a split-plot design was established. The design includes one quantitative variable for nitrogen addition (i.e. gradient approach with seven levels applied across plots; referred to as blocks in @halbritter2025data) and two factorial variables for warming (2 levels) and clipping/grazing (4 levels). Nine experimental plots containing eight 50 x 50 cm subplots were established at each of the two origin sites (alpine and sub-alpine, n = 144) (@fig-methods). At the boreal destination site, plots and subplots were prepared to receive transplanted subplots (see below). Six of the subplots in each plot were fenced in early spring 2020, and two remained outside for the natural grazing treatment. Each subplot received a combination of warming, nitrogen addition and clipping/grazing treatments in two steps. First, nitrogen treatments were randomly assigned at the plot level, to avoid nitrogen contamination between the subplots within a plot. Second, the warming and grazing treatments were randomised within each plot, with the two naturally grazed subplot located outside the fenced areas.

At the end of the growing season in 2019, the subplots assigned to the warming treatment were transplanted from the origin (alpine and sub-alpine) to the destination (sub-alpine and boreal) sites, respectively, simulating an approximately 3.0 °C warmer growing season temperature (Figure S1, Table S1). Subplots to be transplanted were cut to a depth of 5 - 25 cm and the entire community, including roots and soils, was excavated and transplanted. The ambient climate subplots were not excavated, because studies from similar systems have shown that transplanting does not affect species composition or community dynamics [@Yang2018-uq; @Vandvik2020-hu]. If necessary, loose soil under the transplanted subplot was carefully removed, or local soil was added to ensure the soil surface was in plane with the surrounding vegetation.

The nitrogen treatment was applied as a gradient design with seven levels: 0, 0.5, 1, 5, 10, 50, and 100 kg N ha^-1^ y^-1^, and three replicates for the unfertilized controls, for a total of nine plots. We applied a wide range of nitrogen levels to simulate nitrogen addition close to and well beyond the current atmospheric nitrogen deposition in the mountains in south-western Norway [2-3 kg N ha^-1^ y^-1^; @aas2023monitoring] and the critical load for change in species composition in alpine grasslands [c. 5-10 kg N ha^-1^ y^-1^; @bobbink2011review]. Our understanding of environmentally extreme conditions is limited, and such simulations can provide a mechanistic understanding of thresholds for community change [@Kayler2015-pr]. Nitrogen was added to each plot using slow dissolving fertilizer pellets (YaraBela OPTI-NS 27-0-0 (4S)). We used oxidised nitrogen (NO~x~) formed mainly by combustion processes [@holland2005nitrogen], which are the main sources of atmospheric nitrogen deposition in remote regions (i.e., away from intensive agriculture and other sources or reduced nitrogen). The nitrogen was added once at the start and once in the middle each growing season, except in 2020 at the boreal site, where the entire fertilizer dose was applied at the start of the growing season.

Finally, the four clipping or grazing treatments were applied as follows per subplot in each plot: Outside the fence, the naturally grazed treatments were subject to selective consumption by domestic and wild animals, as well nutrient addition and trampling. Inside the fence, were unclipped and ungrazed controls and to isolate the effects of biomass removal, the intermediate and intensive clipping treatments. The vegetation was manually clipped using scissors, at about 2-3 cm above the ground, applied two (intermediate) or four (intensive) times during each growing season. In 2020, the intensive clipped subplots at the alpine site were only cut three times because it was a short growing season due to late snowmelt; an additional cut at the end of the growing season would not have yielded any biomass. In 2021 and 2022, we did not cut Carex species before the species composition was surveyed in early August (first and second cut), because they are difficult to identify without leaf tips.

### Data collection

Microclimate data was recorded continuously every 15 minutes, between July 2019 and September 2022, using Tomst TM4 loggers [@Wild2019-nk]. Air, ground and soil temperature as well as soil moisture were measured at 15, 0 cm above and -8 cm below the ground in all warming and grazing treatments inside the fence and at 0, 1, 10, and 100 kg N ha^-1^ y^-1^ (n = 72 loggers). At the alpine site, the plot with 0.5 kg N ha^-1^ y^-1^ also received climate loggers, because this block is located further away from the rest and we wanted to observe any differences in the microclimate. The warming treatment increased air, ground, and soil temperature by c. 3 °C and decreased soil moisture (Figures S1, S2, Table S1,S3), whereas nitrogen addition and grazing had no effect on the microclimate (Figure S2, Table S3).

After three years, we recorded plant species composition and vegetation height for all subplots at peak growing season. We visually estimated the percentage cover of each vascular plant species to the nearest 1 % on the 25 x 25 cm inner area of each subplot. The total coverage in each subplot can exceed 100 %, due to layering of the vegetation. For species identification we followed the nomenclature in @Lid2010-ck. The vegetation height was measured at four points in each subplot. The community biomass of all unclipped and ungrazed subplots was harvested 2-3 cm above the ground at peak growing season, dried at 65 °C for 72 h, and weighed.

We characterized the environmental preference of the plant communities, using Ellenberg indicator values for light, temperature, nutrients, reaction, and moisture as species' environmental affinities, extracted from a database for European vascular plants [@tichy2023ellenberg]. Higher indicator values indicate species' associations with environments with higher light, temperature, nutrients, alkalinity, and soil moisture. Note that the indicator value for nutrients was initially defined for soil nitrogen availability, but is now often used more generally for nutrients [@ellenberg1992zeigerwerte]. In addition, we used a grazing indicator value, where higher values indicate tolerance for higher severity of grazing, developed for European plants [@midolo2023disturbance]. Species' environmental affinities are available for 87 % of the species (47 out of 54 species) covering on average 70.5 ± 0.72 % of the total cover of the subplots.

### Biomass estimation

Community biomass was only collected in the unclipped and ungrazed subplots (n = `r targets::tar_read("n_calibration_biomass_plots")`). We used these data to estimate community biomass for all subplots in 2022, after 3 years of treatment. We first ran a linear regression for the subplots with collected community biomass as response and summed vascular plant cover x vegetation height and the main and interactive effects of warming and log transformed nitrogen addition as predictors. Then the AIC's from models with all combinations of these predictors were compared to find the most parsimonious model. The model with the lowest AIC contained vascular plant cover x height and log transformed nitrogen addition (R^2^ = `r round(standing_biomass_model_output$r.squared, 2)`), suggesting that nitrogen affected the relationship between vegetation volume and biomass; specifically we found denser vegetation in plots with more fertilisation (Figure S3, Table S4, S5). The regression coefficients from this model were then used to estimate community biomass (g m^-2^) from vascular plant cover x height in all subplots.

### Statistical analysis

All analyses were run separately for the two origin sites (alpine and sub-alpine communities) to assess whether similar mechanisms operate for the two plant communities. The main analyses were run for the clipping treatments, because we were mostly interested in the effect of biomass removal. We repeated some of the analysis for the grazing treatment to generalise the results to a natural grazing regime.

We first conducted more traditional analyses of the effects of the three global change treatments on both community biomass and community diversity after 3 years of treatment to evaluate the main and interactive effects of warming, nitrogen addition, and clipping. We assessed three diversity indices: species richness (n; number of species), Shannon diversity $H' = -\Sigma_{i} p_{i} * ln(p_{i})$, where $p_{i}$ is the proportional cover of species ${i}$, and species evenness $\frac{H'}{log(n)}$ using the vegan package [@oksanen2022vegan].

We used a fixed effects linear model with a normal error distribution. The experimental design did not allow running a mixed-effects model, because of the gradient design for the nitrogen treatment. Clipping was transformed to a numerical scale with 0 (unclipped), 2 (intermediate) or 4 (intensive) cuts per growing season. The nitrogen levels ranged from low to very high levels and were distributed on a logarithmic scale. To assess the shape of the response to nitrogen addition, we tested three models (including the main and interactive effects of all drivers) that reflected different assumptions about this shape: a) a linear term (N), b) a logarithmic term (log(N)) to account for the distribution of the nitrogen levels, and c) a quadratic term (N^2^) allowing for a concave/convex relationship. To find the most parsimonious model, we compared the AIC values of all three models and chose the one with the lowest AIC values. If the models differed in 𝚫AIC ≤ 2, we chose the simplest model (i.e., the fewest interaction terms). We preferred the logarithmic before the linear model, because it reflects a more realistic assumption.

The most parsimonious model for community biomass was the logarithmic model. The most parsimonious model for all three diversity indices was the linear model, except for a quadratic model for Shannon diversity at the sub-alpine site. For consistency, and because 𝚫AIC between the quadratic and linear model was relatively low (𝚫AIC = 4.1), we used the linear model for all diversity indices.

To test our primary prediction that warming, nitrogen addition, and clipping affect plant diversity indirectly via community biomass rather than directly, we ran separate structural equation models (SEM) for each site (@fig-concept b; Equation 1). In these models, nitrogen addition was log transformed and clipping was coded as a numeric variable (see above). We ran separate SEMs for Shannon diversity, species richness, and evenness. We also repeated the SEMs described above for natural grazing (as a factor with two levels ungrazed and grazed) instead of the clipping treatment. All SEMs were run using the piecewiseSEM package [@lefcheck2016piecewisesem].

```{r}
#| label: eq-sem-modell
#| echo: true
#| eval: false

model <- psem(
  lm(diversity ~ biomass + warming + nitrogen + clipping, dat), # direct effects
  lm(biomass ~ warming + nitrogen + clipping, dat) # indirect effects via biomass
  )
```

[Equation 1: Structural equation model used to test the direct effects of warming, nitrogen addition and clipping on diversity indices (e.g.Shannon diversity) and indirectly via community biomass.]{style="color: #808080; font-size: 0.9em;"}

To understand the underlying mechanisms of how the three global change drivers affect the plant community composition, we described changes in the environmental niches of these communities in response to the treatments using Ellenberg indicator values (light, temperature, nutrients, reaction, and moisture) and one grazing disturbance indicator value (*hereafter* environmental affinities). We calculated the community weighted mean of each environmental affinity using the traitstrap package [@telford2021traitstrap; @maitner2023bootstrapping]. We then tested how warming, nitrogen addition (log transformed), and clipping affected each of the affinities. Separate linear models were run for each global change treatment with a normal error distribution.