---
editor: visual
---

```{r}
#| label: set-up-methods
#| echo: false
#| message: false
library(targets)
library(tarchetypes)

tar_load(standing_biomass_model_output)

```

## Methods

### Study area and experimental design

Three semi-natural, species rich, perennial grassland sites (boreal, sub-alpine, and alpine) were established along an elevational gradient in south-western Norway (@fig-methods; 7.17°E, 60.88°N). The lowest site was used as warming treatment for the transplanted sub-alpine communities. The sites were approximately 400 m apart in elevation (469, 920, and 1290 m a.s.l.), corresponding to a temperature difference of approximately 3°C in summer air, ground and soil temperature between the sites over the duration of the experiment (air: 3.09 ± 0.186 °C, ground: 3.06 ± 0.176 °C, and soil: 2.82 ± 0.142 °C; Supplementary Figure S1 and Table S2). Annual precipitation was similar at the two lowest sites and approximately 800 mm higher at the alpine site (Supplementary Table S2). The sites were located on calcareous bedrock with loamy sand and grazed by sheep, goats and deer at the boreal and sub-alpine sites, and mainly sheep and reindeer at the alpine site [@halbritter2025data]. Annual plant productivity is similar at the boreal and sub-alpine site and about four times lower at the alpine site, while species richness is approximately twice as high at the alpine site compared to the sub-alpine site (Supplementary Table S2). Grazing pressure is 3 - 5 times higher in the sub-alpine site compared to the boreal and alpine site [@halbritter2025effects]. Other factors such as land-use history, vegetation type, slope and exposure were kept as constant as possible [@halbritter2025data]. To simulate the natural grazing regime at each site, and to avoid accumulation of biomass, the biomass within the fenced experimental area was removed at the end of each field season.

![Experimental design and map of location. An experiment using a split-plot design to assess warming, nitrogen addition, clipping and grazing replicated in two sites, an alpine and a sub-alpine plant community. Warmed communities were transplanted from the alpine to the sub-alpine or from the sub-alpine to the boreal site, in combination with nitrogen addition, clipping and grazing treatments. The treatments are ambient (gray), warming (red), nitrogen deposition (gray to green under ambient temperature and red to brown under warmed conditions, crossed with four levels of clipping and grazing (yellow); none (C), intermediate clipping (one scissor), intensive clipping (two scissors) and natural grazing (sheep). The arrows indicate the direction of transplant (warming treatment). Note that the nitrogen levels were applied at the plot level and the warming and clipping/grazing treatments were applied at the subplot level. Inset shows location of the study sites in south-western Norway.](figures/methods.jpg){#fig-methods width="800"}

In summer 2019, a global change experiment using a split-plot design was established. The design includes one quantitative variable for nitrogen addition with seven different levels (gradient approach applied across plots; referred to as block in @halbritter2025data) and two factorial variables for warming (2 levels, randomized for subplots within plots for a total of 40 plots/treatment/site level) and grazing/clipping (4 levels, randomized for subplots within each warming treatment and plot for a total of 20 plots/treatment/site level). Ten experimental plots containing eight 50 x 50 cm subplots were established at the two origin sites (alpine and sub-alpine, n = 160) (@fig-methods) and at the boreal destination site, plots and subplots were prepared to receive transplanted subplots (see below). Six of the subplots in each plot were fenced in early spring 2020, and two remained outside to experience natural grazing. Each subplot received a combination of warming, nitrogen addition and grazing treatments in two steps. First, each plot was randomly assigned a nitrogen level, to avoid nitrogen contamination between the subplots within the plots. Second, the warming and grazing treatments were randomized within each plot, with the exception of the naturally grazed subplots, that were located outside the fenced area.

At the end of the growing season in 2019, half of the subplots in each plot were transplanted from the origin (alpine and sub-alpine) to the destination (sub-alpine and boreal) sites, respectively, simulating an approximately 3.0 °C warmer growing season temperature (Supplementary Figure S1, Table S1). Subplots to be transplanted were cut to a depth of 5 - 25 cm and the entire community, including roots and soils, was removed and transplanted. The ambient subplots were not excavated, because studies from similar systems have shown that transplanting does not affect species composition [@Yang2018-uq; @Vandvik2020-hu]. If necessary, loose soil was carefully removed from the underside of the subplot, or local soil was added to the gap or around the edges to ensure the soil surface was in plane with the surrounding vegetation.

Each plot received one of the following seven nitrogen treatments: 0, 0.5, 1, 5, 10, 50, and 100 kg N ha^-1^ y^-1^, with three of these plots at a site serving as unfertilized controls and received no nitrogen addition. We applied a large range of nitrogen levels to simulate nitrogen addition close to and well beyond the current atmospheric nitrogen deposition in the mountains in south-western Norway [2-3 kg N ha^-1^ y^-1^; @aas2023monitoring] and the critical load for change in species composition in alpine grasslands [c. 5-10 kg N ha^-1^ y^-1^; @bobbink2011review]. Our understanding of environmentally extreme conditions is limited, and such simulations can provide a mechanistic understanding of thresholds for community change [@Kayler2015-pr]. Nitrogen was added to each plot using slow dissolving fertilizer pellets (YaraBela OPTI-NS 27-0-0 (4S)). We used oxidised nitrogen (NO~x~) formed mainly by combustion processes [@holland2005nitrogen], which are the main sources of atmospheric nitrogen deposition in remote regions (i.e., away from intensive agriculture and other sources or reduced nitrogen). The nitrogen was added once at the start and once in the middle of the growing season from 2020 - 2022, except in 2020 at the boreal site, where the whole dose of fertilizer was applied at the start of the season.

Finally, each subplot was assigned one of four clipping or grazing treatments: unclipped and ungrazed control, intermediate clipping, intensive clipping (all inside grazing exclosure), or natural grazing (outside grazing exclosure). We manually clipped the vegetation using scissors. The intermediate and intensive clipped subplots were cut two or four times about 2-3 cm above the ground during each growing season. The subplots located outside the fence were naturally grazed by domestic and wild animals and therefore also experiencing selectivity of grazers and nutrients addition and trampling of grazers. The clipping treatment was included to eliminate such effects of grazing and isolate the effects of biomass removal. In 2020, the intensive clipped subplots at the alpine site were only cut three times because it was a short growing season due to late snowmelt. An additional cut at the end of the growing season would not have yielded any biomass. In 2021 and 2022, we did not cut Carex species before the species composition was surveyed in early August (first and second cut), because they are difficult to identify without leaf tips.

For more details on the site selection, experimental setup, and methodology see @halbritter2025data.

### Data collection

Microclimate data was recorded continuously every 15 minutes, between July 2019 and September 2022, using Tomst TM4 loggers [@Wild2019-nk]. Air, ground and soil temperature as well as soil moisture were measured at 15, 0 cm above and -8 cm below the ground in all warming and grazing treatments inside the fence and at 0, 1, 10, and 100 kg N ha^-1^ y^-1^ (n = 72 loggers). At the alpine site, the plot with 0.5 kg N ha^-1^ y^-1^ also received climate loggers, because this block is located further away from the rest and we wanted to make sure to catch any difference in the microclimate. The warming treatment increased air, ground and soil temperature by c. 3 °C and decreased soil moisture (Supplementary Figures S1 and S2, Table S1 and S3), but nitrogen addition and grazing had no effect on the microclimate (Supplementary Figure S2 and Table S3).

Plant species composition and vegetation height were recorded at peak growing season in all subplots before any of the treatments in 2019 and after 3 years in 2022. We visually estimated the percentage cover of each vascular plant species to the nearest 1 % on the 25 x 25 cm inner area of each subplot. The total coverage in each subplot can exceed 100 %, due to layering of the vegetation. For species identification we followed the nomenclature from @Lid2010-ck. The average vegetation height was measured at four points in each plot.

In 2022, community biomass was harvested at peak growing season in all unclipped and ungrazed subplots. The biomass was harvested 2-3 cm above the ground, dried at 65 °C for 72 h, and weighed.

We characterized the environmental preference of the plant communities, using Ellenberg indicator values for light, temperature, nutrient, reaction, and moisture as environmental affinities, extracted from a database for European vascular plants [@tichy2023ellenberg]. Higher indicator values for a species are associated with distributions in environments with higher light, temperature, nutrients, alkalinity, and soil moisture. Note that the indicator value for nutrients was initially defined for nitrogen content, but is now more often used generally for nutrients [i.e., nitrogen and phosphorus; @ellenberg1992zeigerwerte]. In addition, we used a grazing disturbance indicator value (i.e. higher values indicate tolerance for higher severity of grazing) developed for European plants [@midolo2023disturbance]. These affinities are available for 87 % of the species (47 out of 54 species) covering on average 70.5 ± 0.716 % of the total cover of the subplots.

### Biomass estimation

Community biomass was only collected in the unclipped and ungrazed subplots in 2022 (n = 40). We used these data to estimate community biomass for all subplots pre-treatment (2019) and after 3 years of treatment (2022). We first ran a linear regression for the subplots with collected community biomass as response and summed vascular plant cover x vegetation height and the main and interactive effects of warming and log transformed nitrogen addition as predictors. Then the AIC's from models with all combinations of these predictors were compared to find the most parsimonious model. The model with the lowest AIC contained vascular plant cover x height and log transformed nitrogen addition (R^2^ = `r round(standing_biomass_model_output$r.squared, 2)`), suggesting that nitrogen affected the relationship between vegetation volume and biomass (i.e. denser vegetation in plots with more fertilization; Supplementary Figure 3, Table S4, and S5). The regression coefficients from this model was then used to estimate community biomass (g m^-2^) from vascular plant cover x height in all subplots.

### Statistical analysis

All analyses were run separately for the two origin sites (alpine and sub-alpine communities) to assess whether similar mechanisms operate for the two plant communities. The main analyses were run for the clipping treatments, because we were mostly interested in the effect of biomass removal. We repeated some of the analysis for the grazing treatment as well to generalize the results to grazing.

We first conducted more traditional analyses of the effects of the three global change treatments on both community biomass and community diversity after 3 years of treatment to evaluate the main and interactive effects of warming, nitrogen addition and clipping as predictors. We assessed three diversity indices: species richness (n; number of species), Shannon diversity $H' = -\Sigma_{i} p_{i} * ln(p_{i})$, where $p_{i}$ is the proportional abundance of species ${i}$, and species evenness $\frac{H'}{log(n)}$ using the vegan package [@oksanen2022vegan].

We used a fixed effects linear model with a normal error distribution. The experimental design did not allow running a mixed-effects model, because we used a gradient approach for nitrogen addition. Clipping was transformed to a numerical scale with 0 (unclipped), 2 (intermediate) or 4 (intensive) cuts per growing season. The nitrogen levels we chose ranged from low to very high levels and were unevenly distributed on a linear scale. To assess the shape of the response to nitrogen addition, we tested three models (including the main and interactive effects of all drivers) that reflected different assumptions about this shape: a) a linear term (N), b) a logarithmic term (log(N)) to account for the uneven distribution of the nitrogen levels, and c) a quadratic term (N^2^) allowing for a concave/convex relationship. To find the most parsimonious model, we compared the AIC values of all three models and chose the one with the lowest AIC values. If the models differed in 𝚫AIC ≤ 2, we chose the simplest model (i.e., the fewest interaction terms). We preferred the logarithmic before the linear model, because it reflects a more realistic assumption than the linear model.

The most parsimonious model for community biomass was the logarithmic model. The most parsimonious model for all three diversity indices was the linear model, except for a quadratic model for Shannon diversity at the sub-alpine site. For consistency, and because 𝚫AIC between the quadratic and linear model was relatively low (𝚫AIC = 4.1), we used the linear model.

We ran the same analysis for grazing (as a factor with two levels ungrazed and grazed) instead of clipping.

To test our primary prediction that warming, nitrogen addition and clipping affect plant diversity via community biomass rather than directly, we ran separate structural equation models (SEM) for each site (@fig-concept b; Equiation 1). Nitrogen addition was log transformed and clipping was coded as a numeric variable representing the number of times the plot was clipped (see above). We ran separate SEMs for Shannon diversity, species richness, and evenness. We also repeated the SEMs described above for natural grazing (as a factor with two levels ungrazed and grazed) instead of the clipping treatment. All SEMs were run using the piecewiseSEM package [@lefcheck2016piecewisesem].

```{r}
#| label: eq-sem-modell
#| echo: true
#| eval: false

model <- psem(
  lm(diversity ~ biomass + warming + nitrogen + clipping, dat), # direct effects
  lm(biomass ~ warming + nitrogen + clipping, dat) # indirect effects via biomass
  )
```

[Equation 1: Structural equation model used to test the direct effects of warming, nitrogen addition and clipping on Shannon diversity and indirectly via community biomass.]{style="color: #808080; font-size: 0.9em;"}

To understand the underlying mechanisms of how the three global change drivers affect the plant community composition, we described changes in the environmental niches of these communities in response to the treatments using Ellenberg indicator values (light, temperature, nutrients, reaction, and moisture) and one grazing disturbance indicator value (hereafter environmental affinities). We calculated the community weighted mean of each environmental affinity using the traitstrap package [@telford2021traitstrap; @maitner2023bootstrapping]. We then tested how the effect of warming, nitrogen addition (log transformed), and clipping affected each of the affinities. Separate linear models were run for each global change treatment with a normal error distribution.