---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false
library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(piecewiseSEM)
library(patchwork)
library(ggpubr)

```

# Results

### Global change drivers affect standing biomass and diversity in complex ways

```{r}
#| label: mean-bio
#| echo: false

tar_load(biomass_div, store = here::here(tar_config_get("store")))

bio <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_bio), 2), 
            se = round(sd(final_bio)/sqrt(n()), 2))

div <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_diversity), 2), 
            se = round(sd(final_diversity)/sqrt(n()), 2))

```

Standing biomass was twice as high at the sub-alpine (`r bio$mean[2]` ± `r bio$se[2]` g m^-2^) compared to the alpine site (`r bio$mean[1]` ± `r bio$se[1]` g m^-2^), and was affected by all three global change drivers, warming, nitrogen addition and clipping (@fig-1-biomass-diversity a). At both sites, warming and nitrogen addition increased standing biomass significantly, while clipping reduced standing biomass but more so with warming (WxC interaction). Shannon diversity was 1.5 times higher at the alpine (`r div$mean[1]` ± `r div$se[1]`) compared to the sub-alpine site (`r div$mean[2]` ± `r div$se[2]`) and was only affected by nitrogen addition. Diversity decreased with increasing nitrogen at both sites, in particular at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity c).

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: Standing biomass (g m^-2^) (a) and Shannon diversity (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (grey) and warm conditions (pink) and the shapes show the different grazing treatments (circle = control, square = medium, triangle = intensive grazing). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters within panels indicate significant (black) model terms (W = warming, N = nitrogen addition, C = cutting, S = site, combined letters, e.g. WxC = interactions).

tar_read(bio_div_figure, store = here::here(tar_config_get("store")))

```

Overall, the relationship between standing biomass and Shannon diversity was negative (F<sub>1, 142</sub> = 53.88, *P* ≤ 0.001) and differed between the two sites and under ambient and warm climate (@fig-bio-div; @tbl-bio-div). At the alpine site, diversity decreases with increasing biomass under warming, but not under ambient temperature. Contrary, in the sub-alpine site, there was a negative relationship between standing biomass and diversity under ambient conditions, however not with warming. This suggests that the these drivers affect diversity through biomass only under specific conditions and the relationship differ between the sites.

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 5
#| fig-cap: Relationship of change in standing biomass and Shannon diversity. The colors indicate ambient (grey) and warm conditions (pink) and the shapes show the different grazing treatments (circle = control, square = medium, triangle = intensive grazing, diamond = natural grazing). Filled symbols are from the alpine site and open symbols from the sub-alpine site. The lines show the prediction line of the model for the alpine (solid) and sub-alpine site (dashed).

tar_read(standingB_div_final_figure, store = here::here(tar_config_get("store")))

```

```{r}
#| label: tbl-bio-div
#| echo: false
#| eval: false
#| warning: false
#| tbl-cap: Model output for the relationship between standing biomass and Shannon diversity.

tar_load(standingB_div_final_model, store = here::here(tar_config_get("store")))

tidy(standingB_div_final_model) |>
  mutate(term = case_match(term,
                           "(Intercept)" ~ "Intercept",
                           "log(final_bio)" ~ "Log(Biomass)",
                           "origSiteIDSub-alpine" ~ "Site",
                           "warmingWarming" ~ "Warming",
                           "log(final_bio):origSiteIDSub-alpine" ~ "Log(Biomass)xSite",
                           "log(final_bio):warmingWarming" ~ "Log(Biomass)xWarming",
                           "origSiteIDSub-alpine:warmingWarming" ~ "SitexWarming",
                           "log(final_bio):origSiteIDSub-alpine:warmingWarming" ~ "Log(Biomass)xSitexWarming")) |>  
  select("Term" = term, "Estimate" = estimate, `Std. Error` = std.error, `t value` = statistic, `P value` = p.value) |>
  gt() |> 
  gt::fmt_number(columns = c(Estimate, `Std. Error`, `t value`), decimals = 2) |> 
  gt::fmt_number(columns = c(`P value`), decimals = 3) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_column_labels(columns = c(Term, Estimate, `Std. Error`, `t value`, `P value`))) |> 
    tab_style(
          style = list(
            cell_text(weight = "bold")
          ),
          locations = cells_body(
            columns = c(Term, Estimate, `Std. Error`, `t value`, `P value`),
            rows = `P value` <= 0.05
          ))

```

When testing the effect of the three global change drivers separately for each sites, we found that standing biomass was co-limited by warming and nitrogen addition at the alpine site, with an additive effect for the interaction of the two drivers (WxN interaction; @fig-1-biomass-diversity b). In addition, there were interactive effects, where clipping decreased standing biomass more with warming (WxC interaction) and with nitrogen addition (NxC interaction). Standing biomass increased with warming and nitrogen addition, and decreased with clipping at the sub-alpine site, but we found no interactive effects. At the alpine site, the combination of warming and nitrogen addition increased Shannon diversity when little nitrogen was added, but decreased diversity at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity d). In addition, clipping could mitigate the negative effects of warming and nitrogen addition on diversity at the higher levels of nitrogen addition (WxNxC interaction). At the sub-alpine site, warming and nitrogen addition had a negative effect on diversity and again increasingly at higher nitrogen levels. There was no effect of clipping or interactive effects at the sub-alpine site.

We found similar patterns for the impact of the three global change drivers on species richness and evenness (Supplemenatry Figure S5). Species richness at the alpine site decreased with warming and nitrogen addition, but clipping had no effect. Evenness decreased with nitrogen addition in the sub-alpine site, but warming had no effect.

### Do global change drivers operate on diversity through the same mechanisms?

We expected warming, nitrogen addition and clipping to operate on Shannon diversity through the same mechanisms, namely standing biomass (H1). Warming and nitrogen addition had the expected positive effect on standing biomass, while clipping had a negative effect (@fig-cut-final-diversity a and b; Supplementary Table S4). The three global change drivers operated on diversity through standing biomass only at the alpine site and increased standing biomass resulted in a negative net effect on diversity. Interestingly, warming also had a direct positive effect on diversity. Contrary, at the sub-alpine site, the three global change drivers had the same effect on standing biomass, but did not indirectly affect diversity. However, warming and nitrogen addition had a direct negative effect on diversity.

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and cutting and indirectly through biomass on Shannon diversity a) for the alpine and b) sub-alpine site. The numbers indicate the regression coefficients, the linetype shows the strength (solid = significant, thick dashed = trend, and thin dashed = non-significant), and the colour the direction of the effect (blue = positive, red = negative). 

tar_load(cut_final_diversity, store = here::here(tar_config_get("store")))

cut_final_diversity[[1]]

```

The patterns were very similar for the other two diversity indices (Supplementary Figures S6 and S7). There was a large indirect effect via standing biomass on species richness at the alpine site, but it was not significant. In addition, there was a trend for a direct effect of nitrogen addition on richness. Evenness had a direct effect of cutting on evenness instead of warming at the sub-alpine site.

*(Maybe drop the grazing, since it does not show very similar patterns?):* At alpine site no indirect effects via biomass, at sub-alpine site warming has no direct effect on biomass (Supplementary Figures S8-10).

### Trait responses to global change of winners and losers

At the alpine site, the plant community had more winners (i.e. species that colonized or increased in cover) and losers (i.e. species that go extinct or decrease in cover) in response to warming, nitrogen addition or both drivers compared to the sub-alpine site (@fig-traits). In response to warming loser species at the alpine site were shorter and had lower temperature requirements. Contrary, winner species in response to warming or nitrogen addition were taller and in response to warming also warm adapted.

At the sub-alpine site, winner species in response to warming were taller. Loser species also responded to clipping at the sub-alpine site (Supplementary Figure S5). Loser species were smaller in response to clipping. Under ambient climate, loser species had lower temperature requirements when they were clipped, but were warmer adapted species under warming and clipping.

### Warming

```{r}
#| label: fig-traits-warming
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Trait distributions for ambient and warmed plots at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values. The colours indicate the ambient (grey) and warming (pink) treatments.

tar_read(traits_warming_plot)

```

### Nitrogen

```{r}
#| label: fig-traits-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Trait distributions for different levels of nitrogen addition at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values. The colours indicate the different nitrogen levels (some gradient).

tar_read(traits_nitrogen_plot)

```

### Clipping

```{r}
#| label: fig-clipping-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Trait distributions for unclipped (control), medium and intensive clipped plots at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values. The colours indicate the control (grey), medium (...), and intensive (...) treatments.

tar_read(traits_clipping_plot)

```

### Biomass

```{r}
#| label: fig-biomass-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: Trait distributions for plots with different standing biomass at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values. The colours indicate different levels of standing biomass (some gradient).

tar_read(traits_biomass_plot)

```

###