---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false
# Set working directory to project root for targets
if(!dir.exists("_targets")) {
  setwd("..")
}
library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(piecewiseSEM)
library(patchwork)
library(ggpubr)
library(cowplot)

```

# Results

### Global change drivers affect community biomass and diversity differently between sites

As expected, at both sites, warming (W) and nitrogen addition (N) significantly increased community biomass, while clipping (C) reduced biomass (@fig-1-biomass-diversity; Supplementary Table S6). At the alpine site, community biomass was co-limited by warming and nitrogen addition (WxN interaction) such that either warming alone or the highest N level alone had similar biomass. The results seemed to be more additive at the sub-alpine site, although the WXN interaction was almost significant (@fig-1-biomass-diversity).

Responses of Shannon diversity to the treatments different considerable at the two sites. At the alpine site, warming increased Shannon diversity over most of the nitrogen addition gradient, but decreased diversity at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity; Supplementary Table S4). Clipping seemed to mitigate this negative effect of warming on diversity at high nitrogen addition (WxNxC). In contrast, at the sub-alpine site, warming and nitrogen addition had a negative effect on diversity throughout the nitrogen addition gradient, although the biggest effect was, again, at higher nitrogen levels (\> 25 kg N ha^-1^ y^-1^ ). There was no effect of clipping or interactive effects at the sub-alpine site.

We found similar patterns for the impact of the three global change drivers on species richness and evenness (Supplementary Figure S4, and Table S4).

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: "Community biomass (g m^-2^) (a) and Shannon diversity (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters (black) within panels indicate significant model terms (W = warming, N = nitrogen addition, C = clipping, S = site, combined letters, e.g. WxC = interactions). Letters in grey indicate almost significant model terms (0.05 < *P* < 0.07)."

tar_read(bio_div_figure)

```

### Do global change drivers operate on diversity and community composition through the same mechanisms?

We tested if warming, nitrogen addition and clipping operate on Shannon diversity through the same mechanisms, namely community biomass, using structural equation models. As we found with the regression models, at both sites, warming and nitrogen addition had the expected positive effect on community biomass, while clipping had a negative effect (@fig-cut-final-diversity a and b; Supplementary Table S7). Also as predicted, community biomass had a negative effect on diversity, albeit weak and non-significant at the sub-alpine site (@fig-cut-final-diversity; @fig-bio-div). However, contrary to our prediction in neither site, all three global change drivers operated on diversity solely indirectly through community biomass. In the alpine, our prediction was largely met, with nitrogen addition and clipping having only indirect effects on diversity, but warming had a substantial positive direct effect on diversity as well as a negative effect mediated through community biomass. In contrast, at the sub-alpine site warming and nitrogen addition had primarily direct negative effects on diversity, while clipping had operated through community biomass.

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping on Shannon diversity, as well as their indirect effects through community biomass for the a) alpine and b) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The colours shows the direction of the effect (black = positive and grey = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, brown = diversity."

tar_load(cut_final_diversity)

cut_final_diversity[[1]]

```

We found similar but weaker patterns for the direct and indirect effects of warming, nitrogen addition, and clipping on species richness and evenness (Supplementary Figures S5, Table S7). For example, the indirect negative effect on species richness via community biomass was weaker and not significant at the alpine site. At the sub-alpine site only warming affected species richness directly while nitrogen addition and clipping had a weaker direct effect on evenness.

<!-- Grazing results: similar but weaker results for grazing instead of clipping. Stronger indirect effect, although not significant, and no direct effect of warming. In the sub-alpine N and G mostly indirect but not significant, while warming had a more direct effect. Need to do stats. -->

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 5
#| fig-cap: "Relationship of change in community biomass and Shannon diversity at the alpine and sub-alpine site. These are the raw data used in the SEMs. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, and triangle = intensive clipping). The size of the symbols indicate the nitrogen addition level. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation."

tar_read(standingB_div_final_figure)

```

To understand the alternative mechanisms of how the three global change drivers directly and indirectly affect plant community composition (H2), we used environmental affinities to describe the changes in environmental niches of these communities. The first principle component (PC) axis of environmental affinities explained 59.0 % of the variation and was associated with a more competitive plant communities (i.e. more temperature and nutrient demanding, as well as shade and grazing tolerant), and partially with higher acidic affinity (reaction; @fig-affinity-pca-sem a). PC axis 2 explained 24.3 % of the variation and correlated with more alkaline and drier communities. The first PC axis separated the alpine and sub-alpine sites as well as the warming treatment at the alpine site. The sub-alpine and the warmed alpine communities were associated with more competitive environmental affinities. Plant communities at both sites that received higher levels of nitrogen and had higher community biomass were located at the lower right corner of the PCA corresponding to higher temperature and nutrient demanding, shade and grazing tolerant and acidic communities (@fig-affinity-pca-sem a and b). The sub-alpine community was not well described by PC axis 1 and separated mostly along PC axis 2 where warmed communities receiving higher levels of nitrogen were associated with more acidic and moist affinities. Since different mechanisms seem to play a role at the alpine and sub-alpine communities, we repeated the PCA using environmental affinities separately by site (PC axis 1 explained 51.2 % and PC axis 2 explained 25 % of the variation; Supplementary Figure S6). This analysis indicates that sub-alpine communities...

```{r}
#| label: affinity-pca
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 10

tar_load(affinity_pca_plot)
affinity_pca_plot

```

```{r}
#| label: fig-affinity-pca-sem
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Multivariate community-weighted mean affinity responses and structural equation models. a) Warming, clipping and nitrogen addition responses and b) biomass responses, determined by a principal component analysis (PCA). The PCA was conducted for the alpine and sub-alpine site together. The plant community originating from the alpine (filled symbols) and sub-alpine (open symbols) sites are shown. The colour represents the warming treatment (black = ambient, red = warming) or biomass (gradient of blue), the shape the clipping treatment (circle = control, square = medium, triangle = intensive clipping), and the point size indicates nitrogen addition level or biomass. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation. Affinities include plant height, temperature, light, moisture, nutrients, reaction, and grazing disturbance. Arrows show trait loadings on the first two principal components. Note that the direct effects of the drivers on community biomass are the same as for Shannon diversity because that part of the SEM analyses is the same. c-f) Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and cutting and indirectly through biomass on affinity PCA axis 1 for the c) alpine and d) sub-alpine site and the affinity PC axis 2 for the e) alpine and f) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The colours shows the direction of the effect (black = positive and grey = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, brown = affinity."

tar_load(cut_affinity)
cut_affinity[[1]]

# Merge the plots using cowplot to preserve existing tags
# Combine the plots using cowplot
#plot_grid(
  #affinity_pca_plot, 
  #cut_affinity[[1]], 
  #ncol = 1, 
  #rel_heights = c(1.1, 1),
  #rel_widths = c(1, 1),
  #align = "v"
#)
```

The three global change drivers operated through similar mechanisms on the PC axis 1 of the environmental affinities compared to the Shannon diversity (@fig-affinity-pca-sem c and d). At the alpine site the effects of nitrogen and clipping on affinity PC axis 1 were mediated through community biomass, while the effect of warming were both direct and indirect (@fig-affinity-pca-sem c). Warming at the alpine site led to more competitive communities including higher temperature and nutrient demanding, more shade and grazing tolerant communities (Supplement Figure 8a). Nitrogen addition in the alpine led to more shade tolerant and acidic communities (Supplement Figure 9a). Grazing had no direct effects on affinity PC axis 1 (Supplement Figure 8c). Indirectly all three drivers affected community biomass and pushed the communities towards more competitive environmental affinities (@fig-biomass-nitrogen).

At the sub-alpine site, warming and nitrogen had mostly indirect effects on the affinity PC axis 1, however there existed a weak effect via community biomass (i.e. the effect was 1.3 times as large as for Shannon diversity; (@fig-affinity-pca-sem d). Warming at the sub-alpine site led to more shade tolerant and moist communities (Supplement Figure 8a), while nitrogen addition shifted communities towards more shade tolerant, acidic and higher nutrient demanding communities (Supplement Figure 8b). Although the indirectly effect of the three drivers mediate through standing biomass was non-significant, higher community biomass at the sub-alpine site led to more shade tolerant, acidic and moist communities (@fig-biomass-nitrogen). Note that higher community biomass at the sub-alpine site affect the affinities less than at the alpine site.

At the alpine site, only nitrogen had a direct effect on affinity PC axis 2 (@fig-affinity-pca-sem e). At the sub-alpine site, non of the direct effect were significant.

When using the affinity PC axis 1 only from the sub-alpine site in the SEM, the global change drivers did still not operate indirectly through community biomass, however, the effect was almost 3 time higher then when using Shannon diversity (Supplementary Figure S7). In addition nitrogen addition affected affinity PC axis 2 directly.

```{r}
#| label: fig-biomass-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Trait distributions for plots with different standing biomass at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values, as well as a mowing and grazing disturbance indicator. The colours indicate different levels of standing biomass."

tar_read(traits_biomass_plot)

```