---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false
library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(piecewiseSEM)
library(patchwork)
library(ggpubr)

```

# Results

### Global change drivers affect standing biomass and diversity in complex ways

```{r}
#| label: mean-bio
#| echo: false

tar_load(biomass_div, store = here::here(tar_config_get("store")))

bio <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_bio), 2), 
            se = round(sd(final_bio)/sqrt(n()), 2))

div <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_diversity), 2), 
            se = round(sd(final_diversity)/sqrt(n()), 2))

```

Standing biomass twice as high at the sub-alpine (`r bio$mean[2]` ± `r bio$se[2]` g m^-2^) compared to the alpine site (`r bio$mean[1]` ± `r bio$se[1]` g m^-2^; @fig-1-biomass-diversity a). Warming and nitrogen addition increased standing biomass significantly at both sites. Clipping reduces standing biomass and more so with warming (WxC interaction).

When doing the analysis separately for each sites, standing biomass increased more with warming and nitrogen addition at the alpine site (WxN; @fig-1-biomass-diversity b). Clipping decreased standing biomass and more so with warming (WxC) and with nitrogen addition (NxC). At the sub-alpine site, standing biomass increased with warming and nitrogen addition, and decreased with cutting, but there were no interactive effects.

Shannon diversity was 1.5 times higher at the alpine (`r div$mean[1]` ± `r div$se[1]`) compared to the sub-alpine site (`r div$mean[2]` ± `r div$se[2]`) and decreased with nitrogen addition, in particular at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity c). When splitting the analysis by origin site, nitrogen addition decrease diversity at both sites (@fig-1-biomass-diversity d). Warming increases diversity at low nitrogen levels in the alpine site and decreases at nitrogen levels above 25 kg N ha^-1^ y^-1^. At the sub-alpine site, warming has a negative effect on the diversity irrespective of the nitrogen levels. At the alpine site, cutting can also mitigate the negative effects of warming and nitrogen addition on diversity (three-way interaction: WxNxC).

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 12
#| fig-cap: Standing biomass (g m^-2^) (a), standing biomass analysed separately by site (b), Shannon diversity (c), and Shannon diversity analysed separately by origin site (d) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (grey) and warm conditions (pink) and the shapes show the different grazing treatments (circle = control, square = medium, triangle = intensive grazing). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters within panels indicate significant (black) model terms (W = warming, N = nitrogen addition, C = cutting, S = site, combined letters, e.g. WxC = interactions).

tar_read(bio_div_figure, store = here::here(tar_config_get("store")))

```


The overall relationship between standing biomass and Shannon diversity is negative (F{1, 142} = 53.88, *P* <= 0.001), but differs between sites and the warming treatment.
In the alpine site, diversity decreases with increasing biomass under warming.
Contrary, in the sub-alpine site, there is a negative relationship between standing biomass and diveresity under ambient conditions.

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 7.5
#| fig-cap: Relationship of change in standing biomass and Shannon diversity.

tar_read(standingB_div_final_figure, store = here::here(tar_config_get("store")))

```

```{r}
#| label: tbl-bio-div
#| echo: false
#| warning: false
#| tbl-cap: Model output for the relationship between standing biomass and Shannon diversity.

tar_load(standingB_div_final_model, store = here::here(tar_config_get("store")))

tidy(standingB_div_final_model) |>
  mutate(term = case_match(term,
                           "(Intercept)" ~ "Intercept",
                           "log(final_bio)" ~ "Log(Biomass)",
                           "origSiteIDSub-alpine" ~ "Site",
                           "warmingWarming" ~ "Warming",
                           "log(final_bio):origSiteIDSub-alpine" ~ "Log(Biomass)xSite",
                           "log(final_bio):warmingWarming" ~ "Log(Biomass)xWarming",
                           "origSiteIDSub-alpine:warmingWarming" ~ "SitexWarming",
                           "log(final_bio):origSiteIDSub-alpine:warmingWarming" ~ "Log(Biomass)xSitexWarming")) |>  
  select("Term" = term, "Estimate" = estimate, `Std. Error` = std.error, `t value` = statistic, `P value` = p.value) |>
  gt() |> 
  gt::fmt_number(columns = c(Estimate, `Std. Error`, `t value`), decimals = 2) |> 
  gt::fmt_number(columns = c(`P value`), decimals = 3) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_column_labels(columns = vars(Term, Estimate, `Std. Error`, `t value`, `P value`)))

```


### Does including standing biomass improve the model?

The best model describing the effect on diversity includes biomass and the main effects of the three drivers. For the sub-alpine site, the best model includes only main effect and not biomass.

```{r}
#| label: var-explained
#| echo: false

tar_load(biomass_div, store = here::here(tar_config_get("store")))

all <- biomass_div |> 
  nest() |>
  mutate(#bio = map(data, ~lm(final_diversity ~ final_bio, data = .)),
         main = map(data, ~lm(final_diversity ~ warming + grazing_num + Namount_kg_ha_y, data = .)),
         biomain = map(data, ~lm(final_diversity ~ final_bio + warming + grazing_num + Namount_kg_ha_y, data = .)),
         #interaction = map(data, ~lm(final_diversity ~ warming:grazing_num + warming:Namount_kg_ha_y + grazing_num:Namount_kg_ha_y + warming:grazing_num:Namount_kg_ha_y, data = .)),
         full = map(data, ~lm(final_diversity ~ warming * grazing_num * Namount_kg_ha_y, data = .))
         #biofull = map(data, ~lm(final_diversity ~ final_bio + warming * grazing_num * Namount_kg_ha_y, data = .))
         ) |>
  mutate(main = map(main, glance),
         biomain = map(biomain, glance),
         full = map(full, glance)) |>
  rename("main effects" = main, "main effects + biomass" = biomain, "main effects + interactions" = full) |> 
  # make long table
        pivot_longer(cols = -c(data),
                     names_to = "model",
                     values_to = "glance") |> 
  mutate(origSiteID = "Both")


site <- biomass_div |> 
  group_by(origSiteID) |>
  nest() |>
  mutate(#bio = map(data, ~lm(final_diversity ~ final_bio, data = .)),
         main = map(data, ~lm(final_diversity ~ warming + grazing_num + Namount_kg_ha_y, data = .)),
         biomain = map(data, ~lm(final_diversity ~ final_bio + warming + grazing_num + Namount_kg_ha_y, data = .)),
         #interaction = map(data, ~lm(final_diversity ~ warming:grazing_num + warming:Namount_kg_ha_y + grazing_num:Namount_kg_ha_y + warming:grazing_num:Namount_kg_ha_y, data = .))
         full = map(data, ~lm(final_diversity ~ warming * grazing_num * Namount_kg_ha_y, data = .)),
         #biofull = map(data, ~lm(final_diversity ~ final_bio + warming * grazing_num * Namount_kg_ha_y, data = .))
         ) |>
  mutate(main = map(main, glance),
         biomain = map(biomain, glance),
         full = map(full, glance)) |>
  rename("main effects" = main, "main effects + biomass" = biomain, "main effects + interactions" = full) |> 
  # make long table
        pivot_longer(cols = -c(origSiteID, data),
                     names_to = "model",
                     values_to = "glance")
site |> 
  unnest(glance) |> 
  select(origSiteID, model, r.squared, adj.r.squared, AIC) |> 
  mutate(r.squared = round(r.squared, 2),
         adj.r.squared = round(adj.r.squared, 2),
         AIC = round(AIC, 2)) |>
  gt()

```

### Does global change operate through the same mechanisms?

We expected warming, nitrogen addition and clipping to operate through the same mechanisms, namely standing biomass to affect diversity (H1). Warming and nitrogen addition had the expected positive effect while cutting had a negative effect on standing biomass (@fig-cut-final-diversity c and d). However, only at the alpine site affected the three global change drivers diversity through standing biomass. Increasing standing biomass resulted in a negative net effect on diversity. At the sub-alpine site, the three drivers had the same effect on standing biomass, but this did not indirectly affect diversity. Warming and nitrogen addition had however a direct negative effect on diversity. Interestingly, warming had a direct positive effect on diversity at the alpine site.

<!-- This is describing panel a and b which we maybe do not want to show? -->

Warming, nitrogen addition and cutting affect change in Shannon diversity indirectly through change in biomass when comparing across the two sites (@fig-cut-final-diversity a). Warming and nitrogen addition have a positive effect and cutting a negative on biomass, with a negative net effect on diversity. Warming also has a direct negative effect on diversity.

When including origin site of the plant communities (site) in the model, the the indirect effect through change in biomass weakens (only a trend; @fig-cut-final-diversity b). Site strongly affects change in biomass and diversity. The effects of the drivers on change in biomass stay similar and nitrogen addition also has a direct negative effect on diversity.

The patterns are the same for species richness and evenness, with some exceptions at the sub-alpine site. Nitrogen addition has only a negative direct trend on species richness Warming does not affect evenness directly in the sub-alpine site, but cutting has a direct negative effect on evenness.

Grazing shows similar patterns as clipping, but the three drivers do not affect any diversity index indirectly via biomass (although the effect size is large). In addition, warming does not affec biomass at the sub-alpine site.

So, why does the increase in biomass not affect diversity at the sub-alpine site?

**Cutting - final standing biomass and diversity**

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and cutting and indirectly through biomass on Shannon diversity a) across sites, b) including site as a moderator, c) for the alpine site only, and d) for the sub-alpine site only. The numbers indicate the regression coefficients, the linetype shows the strength (solid = significant, thick dashed = trend, and thin dashed = non-significant), and the colour the direction of the effect (blue = positive, red = negative). 

tar_load(cut_final_diversity, store = here::here(tar_config_get("store")))

cut_final_diversity[[1]]

```

```{r}
#| label: fig-cut-final-diversity-AW
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and cutting and indirectly through biomass on Shannon diversity a) across sites, b) including site as a moderator, c) for the alpine site only, and d) for the sub-alpine site only. The numbers indicate the regression coefficients, the linetype shows the strength (solid = significant, thick dashed = trend, and thin dashed = non-significant), and the colour the direction of the effect (blue = positive, red = negative). 

tar_load(cut_final_diversity_AW, store = here::here(tar_config_get("store")))

cut_final_diversity_AW[[1]]

```

### Dive into the mechanisms: who is winning and losing?

**Question:** Do winners and losers have different traits?

At the alpine site, with warming cold and dry demanding species are lost. With nitrogen addition, cold demanding species are lost. With warming and nitrogen addition, species that require light and nutrient poor conditions are lost.

With warming, taller, less light, and nutrient demanding species are increasing in cover. Similarly, with nitrogen addition, taller and less light sensitive \$\$species are increasing in cover. With warming and nutrient addition, temperature demanding species are increasing in cover. With increasing temperature taller and temperature demanding species win. With increasing nitrogen, taller plants win, but when the vegetation is cut, this trend is stronger at lower nitrogen levels and smaller at higher nitrogen levels.

At the sub-alpine site with nitrogen addition moisture demanding species lose. With warming and nitrogen addition, cold demanding species lose. With nitrogen addition, taller, colder (?), and nutrient poor species decrease in cover. With warming, taller species increase in cover.

Sub-alpine: When the vegetation is cut, shorter species are lost. With warming and cutting, taller species are lost. With nitrogen addition, taller species increase in cover and more so with cutting in particular at high nitrogen levels. With warming and cutting, light demanding species increase in cover.

WxNxC: with warming and nitrogen drier species increase in cover, but with cutting less so?

```{r}
#| label: fig-traits
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: v

tar_read(traits_figure, store = here::here(tar_config_get("store")))

```