---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false

library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(vegan)
library(patchwork)

```

# Results

### Global change drivers affect biomass and diversity differently between the alpine and sub-alpine community

As expected, community biomass significantly increased with both warming (W) and nitrogen addition (N) in both communities, while clipping (C) reduced biomass (@fig-1-biomass-diversity a; Table S6). In the alpine community, biomass was co-limited by warming and nitrogen addition (WxN interaction; *P = 0.002*; Table S6) such that either warming alone or the highest N level alone had similar biomass and effects of the two treatments were superadditive. In contrast, in the sub-alpine community the effects of warming and nitrogen addition were subadditive, and the effect of warming was similar to a medium level of nitrogen addition (ca. 25 kg N ha^-1^ y^-1^). Finally, the three global change drivers explained more variation in community biomass in the alpine than the sub-alpine community (alpine: R^2^ = 0.80, adj. R^2^ = 0.77; sub-alpine: R^2^ = 0.59, adj. R^2^ = 0.53).

Responses of Shannon diversity to the treatments differed considerably between the two plant communities. In the alpine community, warming increased Shannon diversity over most of the nitrogen addition gradient, but decreased diversity at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity b; Table S6). Clipping partly mitigated this negative effect of warming on diversity at high nitrogen addition (WxNxC; *P = 0.014*; Table S6). In contrast, in the sub-alpine community, warming had a negative effect on diversity throughout the nitrogen addition gradient. High levels of nitrogen addition (\> 25 kg N ha^-1^ y^-1^) in particular decreased diversity in both communities. There was no effect of clipping or interactive effects in the sub-alpine community.

The three global change drivers affected species richness and evenness similarly to diversity (Figure S4, Table S6).

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: "Community biomass (g m^-2^) (a) and Shannon diversity (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) community. The x axis is log transformed with numbers on the original scale for ease of interpretation. The colours indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. Capital letters (black) within panels indicate significant model terms (W = warming, N = nitrogen addition, C = clipping, combined letters, e.g. WxC = interactions). Letters in grey indicate marginally significant model terms (0.05 < *P* < 0.07). R^2^ and Adj. R^2^ are shown in the upper right corner."

tar_read(bio_div_figure)

```

### Do global change drivers operate on diversity and community composition through the same mechanisms?

We tested if warming, nitrogen addition and clipping operate on Shannon diversity through community biomass using SEMs. As we found with the regression models, in both communities, warming and nitrogen addition had the expected positive effect on community biomass, while clipping had a negative effect (@fig-cut-final-diversity a, b; Table S7). Also as predicted, increasing community biomass had a net negative effect on diversity, albeit weak and non-significant in the sub-alpine community (@fig-cut-final-diversity; Figure S4). However, contrary to our prediction, in none of the communities did all three global change drivers impacted diversity solely indirectly through community biomass. In the alpine community, our prediction was largely met, with nitrogen addition and clipping having only indirect effects on diversity, but warming had a substantial positive direct effect on diversity in addition to a negative effect mediated through community biomass. In the sub-alpine community only clipping operated solely indirectly through community biomass. Warming and nitrogen addition had primarily direct negative effects on diversity. Interestingly, warming had the opposite direct effects on diversity in the two communities.

We found similar patterns for the direct and indirect effects of the global change drivers on species richness and evenness (Figures S6, Table S7). The three drivers had a strong indirect negative effect (via biomass) on evenness in the alpine community, while for richness in the alpine community and evenness in the sub-alpine community these indirect effects through biomass were weaker. Warming increased richness and evenness directly in the alpine community. In contrast, in the sub-alpine community, warming had a negative effect on richness. Nitrogen addition and clipping decreased evenness directly in the sub-alpine community.

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping on Shannon diversity, as well as their indirect effects through community biomass for the a) alpine and b) sub-alpine communities. The numbers indicate the regression coefficients, the colour the direction (orange = positive, turquoise = negative),  and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05) and linetype  (solid = significant and dashed = non-significant)."

tar_load(cut_final_diversity)

cut_final_diversity[[1]]

```

The SEMs including grazing showed very similar patterns to those including clipping, but had slightly weaker indirect biomass-mediated effects (Figure S6, Table S8). The strength of the effect of grazing and clipping on community biomass was very similar.

### Shifts in environmental affinities

We first assessed whether species' environmental affinities were consistent with our hypothesis that increased competition for light and consequent loss of high-light demanding species drive lower diversity associated with increasing community biomass. Indeed, affinities shifted to more shade-tolerant species (lower light indicator values) with increasing biomass, as well as with warming and nitrogen addition in both communities (@fig-toenails). Second, we used species' environmental affinities to assess mechanisms other than biomass that could be driving the strong direct effects of warming and nitrogen addition on diversity. We found greater abundance of species with high-temperature and nitrogen affinities in response to warming in the alpine community but not in the sub-alpine community (@fig-toenails). Interestingly, warming shifted the sub-alpine community towards species with high moisture affinities. For the nitrogen addition treatment, we found a slight, but significant decrease in high-N demanding species with increasing N in the sub-alpine community. Greater nitrogen addition also resulted in higher abundance of species with affinities for more acid conditions (reaction indicator) in both communities. With warming and higher biomass, the communities shifted towards species with higher tolerance for grazing in the alpine community (@fig-toenails; Table S9), but none of the affinities showed any response to the grazing treatment (Figure S7, Table S9).

```{r}
#| label: fig-toenails
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Distributions for species' environmental affinities (community weighted Ellenberg indicator values) at different levels of community biomass (a), and for warming (b) and nitrogen addition (c) treatments at the alpine (upper row) and sub-alpine communities (bottom row). Shown are the distributions for environmental affinities light, temperature, nutrient, moisture, salinity and reaction, and grazing disturbance. The colours indicate the mediator and different treatments: a) standing biomass, b) nitrogen addition levels, and c) warming (ambient = black, warming = red). Darker colours indicate higher values of biomass or nitrogen addition. The stars show the level of significance (* for P ≤ 0.05, ** for P ≤ 0.01, *** for P ≤ 0.001)."

# Load the individual plots
tar_load(traits_biomass_plot)
tar_load(traits_warming_plot)
tar_load(traits_nitrogen_plot)

# Combine the plots with tags
(traits_biomass_plot + labs(tag = "a) Community biomass")) +
    (traits_warming_plot + labs(tag = "b) Warming")) +
    (traits_nitrogen_plot + labs(tag = "c) Nitrogen")) +
    plot_layout(ncol = 1, guides = "collect") &
    theme(
        legend.position = "none",
        legend.box = "vertical",
        plot.tag.position = c(0.12, 1.08),
        plot.tag = element_text(size = 12, hjust = 0, vjust = 1),
        plot.margin = margin(t = 10, r = 5, b = 5, l = 5)
    )

```