---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false
library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(piecewiseSEM)
library(patchwork)
library(ggpubr)

```

# Results

### Global change drivers affect standing biomass and diversity differently between sites

```{r}
#| label: mean-bio
#| echo: false

tar_load(biomass_div, store = here::here(tar_config_get("store")))

bio <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_bio), 2), 
            se = round(sd(final_bio)/sqrt(n()), 2))

div <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_diversity), 2), 
            se = round(sd(final_diversity)/sqrt(n()), 2))

```

<!--# Should I mention the main effects only first, since there are interactions in the alpine site -->

Standing biomass was twice as high at the sub-alpine (`r bio$mean[2]` ± `r bio$se[2]` g m^-2^) than the alpine site (`r bio$mean[1]` ± `r bio$se[1]` g m^-2^), and was affected by all three global change drivers, warming, nitrogen addition, and clipping (@fig-1-biomass-diversity; Supplementary Table S4). At both sites, warming (W) and nitrogen addition (N) significantly increased standing biomass, while clipping (C) reduced standing biomass. Standing biomass was co-limited by warming and nitrogen addition at the alpine site (WxN interaction). Clipping decreased standing biomass more with warming (WxC) and with nitrogen addition (NxC). At the sub-alpine site, standing biomass increased with warming and nitrogen addition, and decreased with clipping, but we found no interactive effects.

Shannon diversity was 1.5 times higher at the alpine (`r div$mean[1]` ± `r div$se[1]`) than the sub-alpine site (`r div$mean[2]` ± `r div$se[2]`). At the alpine site, the combination of warming and nitrogen addition increased Shannon diversity when little nitrogen was added, but decreased diversity at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity; Supplementary Table S4). Clipping could mitigate the negative effects of warming and nitrogen addition on diversity at the higher levels of nitrogen addition (WxNxC). At the sub-alpine site, warming and nitrogen addition had a negative effect on diversity and again increasingly at higher nitrogen levels (\> 25 kg N ha^-1^ y^-1^ ). There was no effect of clipping or interactive effects at the sub-alpine site.

We found similar patterns for the impact of the three global change drivers on species richness and evenness (Supplementary Table S4; Supplemenatry Figure S5). Species richness decreased with warming and nitrogen addition at both sites, but clipping had no effect. The same pattern was found for evenness at the alpine site, but clipping could mitigate the negative impacts of warming and nitrogen addition at higher nitrogen levels (WxNxC). At the sub-alpine site, evenness decreased with nitrogen addition, but warming and clipping had no effect.

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: "Standing biomass (g m^-2^) (a) and Shannon diversity (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters (black) within panels indicate significant model terms (W = warming, N = nitrogen addition, C = clipping, S = site, combined letters, e.g. WxC = interactions). Letters in grey indicate almost significant model terms (0.05 < *P* < 0.07)."

tar_read(bio_div_figure, store = here::here(tar_config_get("store")))

```

### Do global change drivers operate on diversity and community composition through the same mechanisms?

To test if warming, nitrogen addition and clipping operate on Shannon diversity through the same mechanisms, namely standing biomass (H1), we ran structural equation models. Warming and nitrogen addition had the expected positive effect on standing biomass, while clipping had a negative effect at both sites (@fig-cut-final-diversity a and b; Supplementary Table S4). The three global change drivers operated on diversity through standing biomass only at the alpine site and increased standing biomass resulted in a negative net effect on diversity. Interestingly, warming also had a direct positive effect on diversity. Contrary, at the sub-alpine site, the three global change drivers had the same effect on standing biomass, but did not indirectly affect diversity. While there seems to be a relationship between standing biomass and Shannon diversity at the sub-alpine site, there is considerable variation (@fig-bio-div) and other processes unrelated to competitive interactions might be more important. Warming and nitrogen addition had a direct negative effect on diversity.

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping, and indirectly through biomass on Shannon diversity for the a) alpine and b) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The linetype and colours shows the direction of the effect (solid = positive and dotted = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, black = diversity."

tar_load(cut_final_diversity)

cut_final_diversity[[1]]

```

We found similar but weaker patterns for species richness and evenness (Supplementary Figures S6). For example, the indirect effect on species richness via standing biomass was there but not significant at the alpine site. In addition, clipping directly affected evenness at the sub-alpine site.

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 5
#| fig-cap: "Relationship of change in standing biomass and Shannon diversity at the alpine and sub-alpine site. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, and triangle = intensive clipping). The size of the symbols indicate the nitrogen addition level."

tar_read(standingB_div_final_figure)

```

To understand the underlying mechanisms of how the three global change drivers affect the plant community composition (H2), we used environmental affinities to describe the environmental niches of these communities. The first principle component (PC) axis explains 62.9 % of the variation and separates the alpine and sub-alpine sites (@fig-affinity-pca). PC axis 1 is associated with a taller and more competitive plant community (i.e. high temperature, nutrient, and lower light demanding community, and grazing tolerant). The warming treatment separates along the first PC axis with warmed communities being more competitive. Higher standing biomass is also associated with PC axis 1. PC axis 2 explains 21% of the variation and is associated with higher reaction (higher pH) and lower moisture.

```{r}
#| label: fig-affinity-pca
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Multivariate community-weighted mean affinity responses to a) warming, clipping and and nitrogen addition or b) biomass, determined by a principal component analysis (PCA). The PCA was conducted for the alpine and sub-alpine site together. The plant community originating from the alpine (filled symbols) and sub-alpine (open symbols) sites are shown. The colour represents the warming treatment (black = ambient, red = warming) or biomass (gradient of blue), the shape the clipping treatment (circle = control, square = medium, triangle = intensive clipping), and the point size indicates nitrogen addition level. Affinities include plant height, temperature, light, moisture, nutrients, reaction, and grazing disturbance. Arrows show trait loadings on the first two principal components."

tar_load(affinity_pca_plot)

affinity_pca_plot

```

When using environmental affinities (PC axis 1 and 2) in the SEM, we find similar patterns as for Shannon diversity (@fig-cut-affinity). We find that the global change drivers affect environmental affinities associated with competitive plant communities (PC axis 1) strongly indirect via standing biomass at the alpine site, while the effect at the sub-alpine site is weak. However, we find stronger indirect effects (although not significant) for PC axis 2. Something why PC axis 2 is stronger, and then explain this is even more evident when the PCA was done separately by site (Supplement XXX).

```{r}
#| label: fig-cut-affinity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and cutting and indirectly through biomass on affinity PCA 1 scores for the a) alpine and b) sub-alpine site and the affinity PCA 2 scores for the c) alpine and d) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The linetype and colours shows the direction of the effect (solid = positive and dotted = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, black = affinity."

tar_load(cut_affinity)

cut_affinity[[1]]

```

??? At the alpine site, all traits except moisture responded to warming. The light requirement and reaction (pH) value decreased with warming, while plant height, nutrient, salinity and temperature increased with warming. Reaction decreased with increasing nitrogen addition. The traits showed similar response to increase in standing biomass as to warming, but moisture also increased and salinity was not significant. No trait in the alpine responded to grazing.

At the sub-alpine site, less traits responded to the global change drivers and biomass. Plant height and moisture increased with warming. Light, reaction (pH), and salinity decreased and temperature increased with nitrogen addition. Plant height decreased with grazing intensity. Light, pH and temperature decreased with increasing standing biomass, while moisture and plant height increased.

```{r}
#| label: fig-biomass-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Trait distributions for plots with different standing biomass at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values, as well as a mowing and grazing disturbance indicator. The colours indicate different levels of standing biomass."

tar_read(traits_biomass_plot)

```