---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false

library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(vegan)

# Load PCA targets for extracting percentages
tar_load(affinity_pca)
tar_load(affinity_alpine_pca)
tar_load(affinity_subalpine_pca)

# Calculate percentages for combined PCA
e_combined <- eigenvals(affinity_pca[[3]])/sum(eigenvals(affinity_pca[[3]]))
pc1_combined <- sprintf("%.1f", round(e_combined[1] * 100, 1))
pc2_combined <- sprintf("%.1f", round(e_combined[2] * 100, 1))

# Calculate percentages for separate site PCAs
e_alpine <- eigenvals(affinity_alpine_pca[[3]])/sum(eigenvals(affinity_alpine_pca[[3]]))
e_subalpine <- eigenvals(affinity_subalpine_pca[[3]])/sum(eigenvals(affinity_subalpine_pca[[3]]))
pc1_alpine <- sprintf("%.1f", round(e_alpine[1] * 100, 1))
pc2_alpine <- sprintf("%.1f", round(e_alpine[2] * 100, 1))
pc1_subalpine <- sprintf("%.1f", round(e_subalpine[1] * 100, 1))
pc2_subalpine <- sprintf("%.1f", round(e_subalpine[2] * 100, 1))

```

# Results

### Global change drivers affect community biomass and diversity differently between sites

As expected, at both sites, warming (W) and nitrogen addition (N) significantly increased community biomass, while clipping (C) reduced biomass (@fig-1-biomass-diversity; Supplementary Table S6). At the alpine site, community biomass was co-limited by warming and nitrogen addition (WxN interaction) such that either warming alone or the highest N level alone had similar biomass. The results seemed to be more additive at the sub-alpine site, although the WXN interaction was almost significant (@fig-1-biomass-diversity).

Responses of Shannon diversity to the treatments different considerable at the two sites. At the alpine site, warming increased Shannon diversity over most of the nitrogen addition gradient, but decreased diversity at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity; Supplementary Table S4). Clipping seemed to mitigate this negative effect of warming on diversity at high nitrogen addition (WxNxC). In contrast, at the sub-alpine site, warming and nitrogen addition had a negative effect on diversity throughout the nitrogen addition gradient, although the biggest effect was, again, at higher nitrogen levels (\> 25 kg N ha^-1^ y^-1^ ). There was no effect of clipping or interactive effects at the sub-alpine site.

We found similar patterns for the impact of the three global change drivers on species richness and evenness (Supplementary Figure S4, and Table S4).

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: "Community biomass (g m^-2^) (a) and Shannon diversity (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters (black) within panels indicate significant model terms (W = warming, N = nitrogen addition, C = clipping, S = site, combined letters, e.g. WxC = interactions). Letters in grey indicate almost significant model terms (0.05 < *P* < 0.07)."

tar_read(bio_div_figure)

```

### Do global change drivers operate on diversity and community composition through the same mechanisms?

We tested if warming, nitrogen addition and clipping operate on Shannon diversity through the same mechanisms, namely community biomass, using structural equation models. As we found with the regression models, at both sites, warming and nitrogen addition had the expected positive effect on community biomass, while clipping had a negative effect (@fig-cut-final-diversity a and b; Supplementary Table S7). Also as predicted, community biomass had a negative effect on diversity, albeit weak and non-significant at the sub-alpine site (@fig-cut-final-diversity; @fig-bio-div). However, contrary to our prediction in neither site, all three global change drivers operated on diversity solely indirectly through community biomass. In the alpine, our prediction was largely met, with nitrogen addition and clipping having only indirect effects on diversity, but warming had a substantial positive direct effect on diversity as well as a negative effect mediated through community biomass. In contrast, at the sub-alpine site warming and nitrogen addition had primarily direct negative effects on diversity, while clipping had operated through community biomass.

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping on Shannon diversity, as well as their indirect effects through community biomass for the a) alpine and b) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The colours shows the direction of the effect (black = positive and grey = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, brown = diversity."

tar_load(cut_final_diversity)

cut_final_diversity[[1]]

```

We found similar but weaker patterns for the direct and indirect effects of warming, nitrogen addition, and clipping on species richness and evenness (Supplementary Figures S5, Table S7). For example, the indirect negative effect on species richness via community biomass was weaker and not significant at the alpine site. At the sub-alpine site only warming affected species richness directly while nitrogen addition and clipping had a weaker direct effect on evenness. Grazing had the same effect on community biomass, diversity, richness and evenness as clipping (Supplementary Figure S6, Table S8). The SEMs including grazing showed similar patterns to the models with clipping and both the direct and indirect effect of the global change drivers via community biomass were the same or weaker.

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 5
#| fig-cap: "Relationship of change in community biomass and Shannon diversity at the alpine and sub-alpine site. These are the raw data used in the SEMs. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, and triangle = intensive clipping). The size of the symbols indicate the nitrogen addition level. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation."

tar_read(standingB_div_final_figure)

```

```{r}
#| label: fig-biomass-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Distributions for affinities for different treatments at the alpine (upper row) and sub-alpine site (bottom row). Shown are the distributions for the Ellenberg indicator values light, temperature, nutrient, moisture, salinity and reaction, and mowing and grazing disturbance. The colours indicate the different treatments: a) warming (ambient = black, warming = red), b) nitrogen addition levels, and c) biomass."

# Load the individual plots
tar_read(traits_biomass_plot)
tar_load(traits_warming_plot)
tar_load(traits_nitrogen_plot)

# Combine the plots with tags
(traits_warming_plot + labs(tag = "a) Warming")) +
  (traits_nitrogen_plot + labs(tag = "b) Nitrogen")) +
  (traits_biomass_plot + labs(tag = "c) Community biomass")) +
  plot_layout(ncol = 1, guides = "collect") &
  theme(legend.position = "none",
        legend.box = "vertical",
        plot.tag.position = c(0.12, 1.08),
        plot.tag = element_text(size = 12, hjust = 0, vjust = 1),
        plot.margin = margin(t = 10, r = 5, b = 5, l = 5))

```