---
editor: visual
---

```{r}
#| label: set-up-res
#| echo: false
#| message: false
library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(piecewiseSEM)
library(patchwork)
library(ggpubr)
library(cowplot)

```

# Results

### Global change drivers affect standing biomass and diversity differently between sites

```{r}
#| label: mean-bio
#| echo: false

tar_load(biomass_div, store = here::here(tar_config_get("store")))

bio <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_bio), 2), 
            se = round(sd(final_bio)/sqrt(n()), 2))

div <- biomass_div |> 
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_diversity), 2), 
            se = round(sd(final_diversity)/sqrt(n()), 2))

```

<!--# Should I mention the main effects only first, since there are interactions in the alpine site -->

Standing biomass was twice as high at the sub-alpine (`r bio$mean[2]` ± `r bio$se[2]` g m^-2^) than the alpine site (`r bio$mean[1]` ± `r bio$se[1]` g m^-2^), and was affected by all three global change drivers, warming, nitrogen addition, and clipping (@fig-1-biomass-diversity; Supplementary Table S4). At both sites, warming (W) and nitrogen addition (N) significantly increased standing biomass, while clipping (C) reduced standing biomass. Standing biomass was co-limited by warming and nitrogen addition at the alpine site (WxN interaction). Clipping decreased standing biomass more with warming (WxC) and with nitrogen addition (NxC). At the sub-alpine site, standing biomass increased with warming and nitrogen addition, and decreased with clipping, but we found no interactive effects.

Shannon diversity was 1.5 times higher at the alpine (`r div$mean[1]` ± `r div$se[1]`) than the sub-alpine site (`r div$mean[2]` ± `r div$se[2]`). At the alpine site, the combination of warming and nitrogen addition increased Shannon diversity when little nitrogen was added, but decreased diversity at nitrogen levels above 25 kg N ha^-1^ y^-1^ (@fig-1-biomass-diversity; Supplementary Table S4). Clipping could mitigate the negative effects of warming and nitrogen addition on diversity at the higher levels of nitrogen addition (WxNxC). At the sub-alpine site, warming and nitrogen addition had a negative effect on diversity and again increasingly at higher nitrogen levels (\> 25 kg N ha^-1^ y^-1^ ). There was no effect of clipping or interactive effects at the sub-alpine site.

We found similar patterns for the impact of the three global change drivers on species richness and evenness (Supplementary Table S4; Supplemenatry Figure S5). Species richness decreased with warming and nitrogen addition at both sites, but clipping had no effect. Similar to Shannon diversity, clipping could mitigate the negative impacts of warming and nitrogen addition at higher nitrogen levels on evenness at the alpine site (WxNxC). At the sub-alpine site, evenness decreased with nitrogen addition, but warming and clipping had no effect.

```{r}
#| label: fig-1-biomass-diversity
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 7.5
#| fig-cap: "Standing biomass (g m^-2^) (a) and Shannon diversity (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters (black) within panels indicate significant model terms (W = warming, N = nitrogen addition, C = clipping, S = site, combined letters, e.g. WxC = interactions). Letters in grey indicate almost significant model terms (0.05 < *P* < 0.07)."

tar_read(bio_div_figure, store = here::here(tar_config_get("store")))

```

### Do global change drivers operate on diversity and community composition through the same mechanisms?

We tested if warming, nitrogen addition and clipping operate on Shannon diversity through the same mechanisms, namely standing biomass (H1), using structural equation models. At both sites, warming and nitrogen addition had the expected positive effect on standing biomass, while clipping had a negative effect at both sites (@fig-cut-final-diversity a and b; Supplementary Table S5). The three global change drivers operated on diversity through standing biomass only at the alpine site and increased standing biomass resulted in a negative net effect on diversity. Interestingly, warming also had a direct positive effect on diversity at the alpine site. Despite seeing a weak relationship between standing biomass and Shannon diversity at the sub-alpine site (@fig-bio-div), the three global change drivers did not indirectly affect diversity via standing biomass. This relationship between standing biomass and Shannon diversity was weak and had considerable variation and these global change drivers might affect diversity through other processes unrelated to competitive interactions. Thus, warming and nitrogen addition had a direct negative effect on diversity.

```{r}
#| label: fig-cut-final-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping, and indirectly through biomass on Shannon diversity for the a) alpine and b) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The colours shows the direction of the effect (black = positive and grey = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, brown = diversity."

tar_load(cut_final_diversity)

cut_final_diversity[[1]]

```

We found similar but weaker patterns for the direct and indirect effects of warming, nitrogen addition, and clipping on species richness and evenness (Supplementary Figures S6). For example, the indirect effect on species richness via standing biomass was weaker and not significant at the alpine site. Also, only warming affected richness directly as well as nitrogen addition and clipping affected evenness directly at the sub-alpine site.

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 5
#| fig-cap: "Relationship of change in standing biomass and Shannon diversity at the alpine and sub-alpine site. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, and triangle = intensive clipping). The size of the symbols indicate the nitrogen addition level. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation."

tar_read(standingB_div_final_figure)

```

To understand the underlying mechanisms of how the three global change drivers directly and indirectly affect plant community composition (H2), we used environmental affinities to describe the changes in environmental niches of these communities. The first principle component (PC) axis of environmental affinities explained 62.9 % of the variation and was associated with a more competitive plant communities (i.e. more temperature and nutrient demanding, as well as shade and grazing tolerant), and partially with higher acidic affinity (reaction; @fig-affinity-pca-sem a). PC axis 2 explained 21% of the variation and correlated with more alkaline and drier communities. The first PC axis separated the alpine and sub-alpine sites as well as the warming treatment at the alpine site. The sub-alpine and the warmed alpine communities were associated with more competitive environmental affinities. Plant communities at both sites that received higher levels of nitrogen and had higher standing biomass were located at the lower right corner of the PCA corresponding to higher temperature and nutrient demanding, shade and grazing tolerant and acidic communities (@fig-affinity-pca-sem a and b). The sub-alpine community separated mostly along PC axis 2 where warmed communities receiving higher levels of nitrogen were associated with more acidic and moist affinities.

```{r}
#| label: fig-affinity-pca
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 10

tar_load(affinity_pca_plot)
affinity_pca_plot

```

```{r}
#| label: fig-affinity-pca-sem
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Multivariate community-weighted mean affinity responses and structural equation models. a) Warming, clipping and nitrogen addition responses and b) biomass responses, determined by a principal component analysis (PCA). The PCA was conducted for the alpine and sub-alpine site together. The plant community originating from the alpine (filled symbols) and sub-alpine (open symbols) sites are shown. The colour represents the warming treatment (black = ambient, red = warming) or biomass (gradient of blue), the shape the clipping treatment (circle = control, square = medium, triangle = intensive clipping), and the point size indicates nitrogen addition level or biomass. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation. Affinities include plant height, temperature, light, moisture, nutrients, reaction, and grazing disturbance. Arrows show trait loadings on the first two principal components. c-f) Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and cutting and indirectly through biomass on affinity PCA axis 1 for the c) alpine and d) sub-alpine site and the affinity PC axis 2 for the e) alpine and f) sub-alpine site. The numbers indicate the regression coefficients and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05). The colours shows the direction of the effect (black = positive and grey = negative). Each driver, mediator and predictor are coloured in a separate colour: red = warming, green = nitrogen, yellow = clipping, blue = biomass, brown = affinity."

tar_load(cut_affinity)
cut_affinity[[1]]

# Merge the plots using cowplot to preserve existing tags
# Combine the plots using cowplot
#plot_grid(
  #affinity_pca_plot, 
  #cut_affinity[[1]], 
  #ncol = 1, 
  #rel_heights = c(1.1, 1),
  #rel_widths = c(1, 1),
  #align = "v"
#)
```

The three global change drivers had similar indirect effects on the PC axis 1 of the environmental affinities as for the Shannon diversity (@fig-affinity-pca-sem c and d). Note that the direct effects of the drivers on standing biomass are the same as for Shannon diversity because that part of the SEM analyses is the same. At the alpine site the effects of nitrogen and clipping on affinity PC axis 1 were mediated through standing biomass, while the effect of warming were both direct and indirect (@fig-affinity-pca-sem c). Warming at the alpine site led to more competitive communities including higher temperature and nutrient demanding, more shade and grazing tolerant communities (Supplement Figure 9a). Nitrogen addition in the alpine led to more shade tolerant and acidic communities (Supplement Figure 9a). Grazing had no direct effects on affinity PC axis 1 (Supplement Figure 9c). Indirectly all three drivers affected standing biomass and pushed the communities towards more competitive affinities (@fig-biomass-nitrogen).

At the sub-alpine site, warming and nitrogen had mostly independent effects on the affinity PC axis 1, however there existed a weak effect via standing biomass (i.e. the effect was 1.5 times as large as for Shannon diversity; (@fig-affinity-pca-sem d). Warming at the sub-alpine site led to more shade tolerant and moist communities (Supplement Figure 9a), while nitrogen addition shifted communities towards more shade tolerant, higher nutrient demanding, and more acidic (Supplement Figure 9b). Although the indirectly effect of the three drivers mediate through standing biomass was non-significant, higher standing biomass at the sub-alpine site led to more shade tolerant, acidic and moist communities. Note that higher standing biomass at the sub-alpine site affect the affinities less than at the alpine site.

At the alpine site, only nitrogen had a direct effect on affinity PC axis 2 (@fig-affinity-pca-sem e). At the sub-alpine site, this direct effect was weaker and not significant. However the direct effect of nitrogen on affinity PCA axis 2 was almost three times as large as on Shannon diversity and was likely related to more acidic communities (Supplement Figure S9b).

```{r}
#| label: fig-biomass-nitrogen
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Trait distributions for plots with different standing biomass at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values, as well as a mowing and grazing disturbance indicator. The colours indicate different levels of standing biomass."

tar_read(traits_biomass_plot)

```