---
format:
    html:
      self-contained: true
    docx: default
    pdf:
      documentclass: scrartcl
      classoption:
        - DIV=11
        - numbers=noendperiod
      papersize: a4
      header-includes:
        - \KOMAoption{captions}{tableheading}
        - \usepackage{placeins}
        - \FloatBarrier
      fig-pos: H
      tbl-pos: H
title: Supporting Information
execute:
  echo: false
bibliography: bibliography.bib
csl: ecology-letters.csl
crossref:
  fig-prefix: "S"
  tbl-prefix: "S"
  fig-title: "Supplementary Figure"
  tbl-title: "Supplementary Table"
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| label: set-up
#| echo: false
#| message: false
# Set working directory to project root for targets
if(!dir.exists("_targets")) {
  setwd("..")
}
library(targets)
library(tarchetypes)
library(tidyverse)
library(lubridate)
library(broom)
library(gt)
library(piecewiseSEM)
library(patchwork)
library(ggpubr)
```

## Environmental data

### Temperature effects on the microclimate

```{r}
#| label: fig-daily-climate
#| fig-height: 6
#| warning: false
#| fig-cap: "a) Daily temperature and soil moisture values, measured over the period of the experiment in the alpine and sub-alpine sites and b) average summer temperature and soil moisture between May and September. Shown are air (30 cm), ground (0 cm), soil temperature (-5 cm), and soil moisture measured with Tomst loggers between August 2019 and October 2022. Colour indicates the ambient (black) and warming (red) treatment. Air, ground and soil temperature were higher and soil moisture was lower in the warming treatment, indicated by the star in each panel."

tar_load(daily_climate_figure)
daily_climate_figure

```


### Stats for temperature effects on the microclimate

```{r}
#| label: tbl-climate-test
#| tbl-cap: "Results from linear regression of the effects of warming on air, ground, and soil temperature and soilmoisture at the alpine and sub-alpine communities. Shown are estimates, standard error, t statistics and P value."


tar_load(microclimate_stats)
microclimate_stats

```

### Environmantal variables for the study sites

```{r}
#| label: method
#| echo: false
#| warning: false

tar_load(daily_temp)

daily_temp2 <- as.data.frame(daily_temp)

temp <- daily_temp2 |>
      mutate(month = month(date),
             year = year(date)) %>%
      filter(month %in% c(5, 6, 7, 8, 9),
             year %in% c(2020, 2021),
             # only controls
             Nlevel %in% c(1, 2, 3),
             grazing == "Control") |> 
  group_by(destSiteID, origSiteID, variable, warming, .drop = TRUE) %>%
  summarise(mean = round(mean(value), 2),
            se = round(sd(value)/sqrt(n()), 1)) |> 
  filter(!c(origSiteID == "Alpine" & warming == "Warming")) |> 
  mutate(siteID = case_when(destSiteID == "Lia" ~ "Alpine",
                            destSiteID == "Joa" ~ "Sub-alpine",
                            destSiteID == "Vik" ~ "Lowland"),
         siteID = factor(siteID, levels = c("Alpine", "Sub-alpine", "Lowland")))

ta <- temp |> 
  filter(variable == "air") |> 
  arrange(siteID)
  

tar_load(consumption)
prod <- consumption |> 
  group_by(siteID) |> 
  summarise(mean = round(mean(Ungrazed), 2),
            se = round(sd(Ungrazed)/sqrt(n()), 1))

cons <- consumption |> 
  group_by(siteID) |> 
  summarise(mean = round(mean(Consumption), 2),
            se = round(sd(Consumption)/sqrt(n()), 1))

tar_load(biomass_div)
div <- biomass_div |> 
  filter(warming == "Ambient", grazing == "Control", Namount_kg_ha_y == 0) |>
  group_by(origSiteID) |> 
  summarise(mean = round(mean(final_richness), 2), 
            se = round(sd(final_richness)/sqrt(n()), 2))

# Create table with site characteristics
# Ensure siteID names are consistent (Lowland -> Boreal)
ta_table <- ta |> 
  mutate(siteID = case_when(
    siteID == "Lowland" ~ "Boreal",
    TRUE ~ as.character(siteID)
  )) |>
  select(siteID, temp_mean = mean, temp_se = se)

prod_table <- prod |> 
  mutate(siteID = case_when(
    siteID == "Lowland" ~ "Boreal",
    TRUE ~ as.character(siteID)
  )) |>
  select(siteID, prod_mean = mean, prod_se = se)

cons_table <- cons |> 
  mutate(siteID = case_when(
    siteID == "Lowland" ~ "Boreal",
    TRUE ~ as.character(siteID)
  )) |>
  select(siteID, cons_mean = mean, cons_se = se)

div_table <- div |>
  mutate(siteID = case_when(
    origSiteID == "Alpine" ~ "Alpine",
    origSiteID == "Sub-alpine" ~ "Sub-alpine",
    TRUE ~ as.character(origSiteID)
  )) |>
  select(siteID, div_mean = mean, div_se = se)

site_table <- data.frame(
  Site = c("Alpine", "Sub-alpine", "Boreal"),
  stringsAsFactors = FALSE
) |>
  # Add air temperature
  left_join(ta_table, by = c("Site" = "siteID")) |>
  # Add precipitation (hardcoded values from text)
  mutate(
    Precipitation_mm = c(2089, 1256, 1292)
  ) |>
  # Add productivity
  left_join(prod_table, by = c("Site" = "siteID")) |>
  # Add consumption
  left_join(cons_table, by = c("Site" = "siteID")) |>
  # Add species richness
  left_join(div_table, by = c("Site" = "siteID")) |>
  # Format values as strings with ± SE
  mutate(
    `Air temperature (°C)` = paste0(temp_mean, " ± ", temp_se),
    `Precipitation (mm)` = as.character(Precipitation_mm),
    `Annual productivity (g m^-2^ y^-1^)` = paste0(prod_mean, " ± ", prod_se),
    `Grazing consumption (g m^-2^ y^-1^)` = paste0(cons_mean, " ± ", cons_se),
    `Species richness` = ifelse(is.na(div_mean), NA_character_, paste0(div_mean, " ± ", div_se))
  ) |>
  # Select only needed columns in desired order
  select(Site, `Air temperature (°C)`, `Precipitation (mm)`,
         `Annual productivity (g m^-2^ y^-1^)`, `Grazing consumption (g m^-2^ y^-1^)`,
         `Species richness`)

```

```{r}
#| label: tbl-site-characteristics
#| echo: false
#| tbl-cap: "Environmental and biotic characteristics of the study sites. Shown are mean ± SE summer air temperature, annual precipitation, annual productivity, grazing consumption and species richness. Air temperature is the average summer temperature (May-September) measured in situ between 2019-2022. Precipitation values are from modeled climate data (2008-2021) by the Norwegian Meteorological Institute. Note that species richness not available for the boreal site, because these plots are transplants from the sub-alpine community."

site_table |>
  gt() |>
  cols_label(
    `Air temperature (°C)` = md("Air temperature (°C)"),
    `Precipitation (mm)` = md("Precipitation (mm)"),
    `Annual productivity (g m^-2^ y^-1^)` = md("Annual productivity (g m<sup>-2</sup> y<sup>-1</sup>)"),
    `Grazing consumption (g m^-2^ y^-1^)` = md("Grazing consumption (g m<sup>-2</sup> y<sup>-1</sup>)"),
    `Species richness` = md("Species richness")
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_column_labels()
  ) |>
  cols_align(align = "left", columns = Site) |>
  cols_align(align = "center", columns = -Site)

```

### Climate and global change drivers

```{r}
#| label: fig-climate
#| warning: false
#| fig-cap: "Global change impact on microclimate. Average summer (June - September) air, ground soil temperature and soil moisture between 2019 and 2022 with increasing nitrogen addition (kg N ha^-1^ y^-1^) at the alpine (upper row) and sub-alpine (lower row) site. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (gray) and warm conditions (pink) and the shapes show the different grazing treatments (circle = control, square = medium, triangle = intensive grazing). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters within panels indicate significant model terms (W = warming)."
tar_load(climate_figure)
climate_figure
```

### Stats for climate and global change drivers

```{r}
#| label: tbl-climate
#| tbl-cap: "Anova table of the effects of warming, nitrogen addition, grazing, and their interactions on air, ground and soil temperature and soilmoisture at the alpine and sub-alpine communities. Shown are Sum of Squares, degrees of freedom (df), F statistics and P value. Significant terms (P < 0.05) are in bold."

tar_load(climate_stats)
climate_stats

```


## Biomass estimation

### Model comparision for estimated community biomass

```{r}
#| label: tbl-biomass-model-comparison
#| tbl-cap: "Model comparison for the relationship between estimated and collected community biomass. Shown are AIC, delta AIC, and R²."

tar_load(biomass_estimation_model_comparison)
biomass_estimation_model_comparison$aic_comparison |>
  gt() |>
  fmt_number(columns = c(AIC), decimals = 1) |>
  fmt_number(columns = c(delta_AIC), decimals = 3) |>
  fmt_number(columns = c(R2), decimals = 3) |>
  cols_label(delta_AIC = "ΔAIC", R2 = "R²") |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_column_labels()
  )

```

### Estimating community biomass

```{r}
#| label: fig-bio-calc-coll
#| echo: false
#| warning: false
#| fig-cap: "Relationship between collected and estimated community biomass in the unfertilized and unclipped control plots. Community biomass is estimated from the community data using *sum of cover x vegetation height*. This estimated biomass is then backtransformed to actual biomass using the biomass collected at peak growing season in the control plots. The colour indicates the warming treatment (ambient = black, warming = red) and the point size the nitrogen addition level. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation. The line type indicate model predictions for 0 (solid) and 100 kg N ha^-1^ y^-1^ (dashed)."


tar_read(standing_biomass_back_fig)

```

### Model output for estimated community biomass

```{r}
#| label: tbl-standing-biomass-model-output
#| echo: false
#| warning: false
#| tbl-cap: "Model output for the relationship between estimated and collected community biomass. Shown are estimates, standard error, t statistics and P value."

tar_load(standing_biomass_model)


tidy(standing_biomass_model) |> 
  mutate(term = case_match(term,
                           "(Intercept)" ~ "Intercept",
                           "biomass_remaining_calc" ~ "Biomass",
                           "Nitrogen_log" ~ "Log(Nitrogen)")) |>  
  select("Term" = term, "Estimate" = estimate, `Std. Error` = std.error, `t value` = statistic, `P value` = p.value) |>
  gt() |> 
  gt::fmt_number(columns = c(Estimate, `Std. Error`, `t value`), decimals = 2) |> 
  gt::fmt_number(columns = c(`P value`), decimals = 3) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_column_labels(columns = c(Term, Estimate, `Std. Error`, `t value`, `P value`))
    )

```


## Biomass-diversity relationship

### Model output for biomass and diversity models

```{r}
#| label: tbl-biomass-diversity-model-output
#| echo: false
#| warning: false
#| tbl-cap: "Model output for the relationship between the global change drivers and community biomass or the diversity indices. Shown are sum of squares, degrees of freedom (df), F statistics and P value. Significant terms (P < 0.05) are in bold."

tar_load(biomass_origin_anova_table)
tar_load(diversity_origin_anova_table)

# Combine the tables
biomass_origin_anova_table |>
  mutate(variable = "Standing biomass") |>
  bind_rows(
    diversity_origin_anova_table |>
      mutate(variable = case_match(diversity_index,
                                   "diversity" ~ "Shannon diversity",
                                   "richness" ~ "Species richness",
                                   "evenness" ~ "Evenness"))
  ) |>
  # Sort variables in desired order
  mutate(variable = factor(variable, levels = c("Standing biomass", "Shannon diversity", "Species richness", "Evenness"))) |>
  # Round values
  mutate(
    sumsq = round(sumsq, 2),
    statistic = round(statistic, 2),
    p.value = if_else(p.value < 0.001, "< 0.001", as.character(round(p.value, 3)))
  ) |>
  # Pivot to wide format to put Alpine and Sub-alpine side by side
  select(variable, origSiteID, term, sumsq, df, statistic, p.value) |>
  pivot_wider(
    names_from = origSiteID,
    values_from = c(sumsq, df, statistic, p.value),
    names_glue = "{origSiteID}_{.value}"
  ) |>
  # Rename columns
  rename(
    "Term" = term,
    "Alpine_Sum of squares" = Alpine_sumsq,
    "Alpine_df" = Alpine_df,
    "Alpine_F" = Alpine_statistic,
    "Alpine_P" = Alpine_p.value,
    "Sub-alpine_Sum of squares" = `Sub-alpine_sumsq`,
    "Sub-alpine_df" = `Sub-alpine_df`,
    "Sub-alpine_F" = `Sub-alpine_statistic`,
    "Sub-alpine_P" = `Sub-alpine_p.value`
  ) |>
  # Reorder columns to put Alpine first
  select(variable, Term, 
         "Alpine_Sum of squares", "Alpine_df", "Alpine_F", "Alpine_P",
         "Sub-alpine_Sum of squares", "Sub-alpine_df", "Sub-alpine_F", "Sub-alpine_P") |>
  # Sort by variable in desired order
  arrange(variable) |>
  # Group by variable
  group_by(variable) |>
  gt() |>
  # Add column spanners for Alpine and Sub-alpine
  tab_spanner(
    label = "Alpine",
    columns = c("Alpine_Sum of squares", "Alpine_df", "Alpine_F", "Alpine_P")
  ) |>
  tab_spanner(
    label = "Sub-alpine", 
    columns = c("Sub-alpine_Sum of squares", "Sub-alpine_df", "Sub-alpine_F", "Sub-alpine_P")
  ) |>
  # Format numbers
  fmt_number(columns = c("Alpine_Sum of squares", "Alpine_F", "Sub-alpine_Sum of squares", "Sub-alpine_F"), decimals = 2) |>
  fmt_number(columns = c("Alpine_df", "Sub-alpine_df"), decimals = 0) |>
  # Make term column always bold
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = "Term")
  ) |>
  # Make Alpine significant values bold
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      columns = c("Alpine_Sum of squares", "Alpine_df", "Alpine_F", "Alpine_P"),
      rows = Alpine_P != "< 0.001" & as.numeric(Alpine_P) <= 0.05
    )
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      columns = c("Alpine_Sum of squares", "Alpine_df", "Alpine_F", "Alpine_P"),
      rows = Alpine_P == "< 0.001"
    )
  ) |>
  # Make Sub-alpine significant values bold
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      columns = c("Sub-alpine_Sum of squares", "Sub-alpine_df", "Sub-alpine_F", "Sub-alpine_P"),
      rows = `Sub-alpine_P` != "< 0.001" & as.numeric(`Sub-alpine_P`) <= 0.05
    )
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      columns = c("Sub-alpine_Sum of squares", "Sub-alpine_df", "Sub-alpine_F", "Sub-alpine_P"),
      rows = `Sub-alpine_P` == "< 0.001"
    )
  ) |>
  # Rename columns at the end to remove prefixes
  cols_label(
    "Alpine_Sum of squares" = "Sum of squares",
    "Alpine_df" = "df",
    "Alpine_F" = "F",
    "Alpine_P" = "P",
    "Sub-alpine_Sum of squares" = "Sum of squares",
    "Sub-alpine_df" = "df",
    "Sub-alpine_F" = "F",
    "Sub-alpine_P" = "P"
  ) |>
  # Set column alignment
  cols_align(
    align = "left",
    columns = "Term"
  ) |>
  cols_align(
    align = "right",
    columns = c("Alpine_Sum of squares", "Alpine_df", "Alpine_F", "Alpine_P",
                "Sub-alpine_Sum of squares", "Sub-alpine_df", "Sub-alpine_F", "Sub-alpine_P")
  )

```

### Global change affects richness and evenness

```{r}
#| label: fig-div-index
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Species richness (a) and evenness (b) with increasing nitrogen addition (kg N ha^-1^ y^-1^) in the alpine (upper row) and sub-alpine (lower row) sites. The x axis is log transformed, the numbers are on the original scale for ease of interpretation. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, triangle = intensive clipping). The points show the raw data, while the lines indicate model predictions with confidence intervals. The capital letters within panels indicate significant (black) model terms (W = warming, N = nitrogen addition, C = clipping, S = site, combined letters, e.g. WxC = interactions). Letters in grey indicate almost significant model terms (0.05 < *P* < 0.07)."

tar_read(div_index_figure)

```

### Relationship of change in community biomass and Shannon diversity

```{r}
#| label: fig-bio-div
#| echo: false
#| warning: false
#| fig-height: 5
#| fig-cap: "Relationship of change in community biomass and Shannon diversity at the alpine and sub-alpine site. These are the raw data used in the SEMs. The colors indicate ambient (black) and warm conditions (red) and the shapes show the different clipping treatments (circle = control, square = medium, and triangle = intensive clipping). The size of the symbols indicate the nitrogen addition level. The nitrogen levels are log transformed, the numbers on the legend are on the original scale for ease of interpretation."

tar_read(standingB_div_final_figure)
```


## Structural equation models

### Statistics for SEMs for Shannon diversity, species richness and evenness at the alpine and sub-alpine site (clipping)

```{r}
#| label: tbl-cut-final-stats2
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 10
#| tbl-cap: "SEM output for direct effects of warming, nitrogen addition, and cutting and indirect effects via community biomass on Shannon diversity, species richness and evenness. Shown are the response, predictors, standard estimate, standard error, degrees of freedome (DF) and P value. Variables in bold have a significant P value."

tar_load(cut_final_diversity)
tar_load(cut_final_richness_evenness)

bind_rows(
  cut_final_diversity[[2]] |> 
    mutate(diversity_index = "diversity",
           Site = Type),
  cut_final_richness_evenness[[2]] |> 
    mutate(diversity_index = case_when(
             str_detect(Type, "richness") ~ "richness",
             str_detect(Type, "evenness") ~ "evenness"
           ),
           Site = case_when(
             str_detect(Type, "Alpine") & !str_detect(Type, "Sub-alpine") ~ "Alpine",
             str_detect(Type, "Sub-alpine") ~ "Sub-alpine"
           ))
) |> 
  mutate(P.Value_num = P.Value,
         is_significant = P.Value <= 0.05) |> 
  mutate(Std.Estimate = round(Std.Estimate, 2),
         Std.Error = round(Std.Error, 2),
         P.Value = if_else(P.Value_num <= 0.001, "≤ 0.001", as.character(round(P.Value_num, 3)))) |> 
  select(Site, Response, Predictor, Std.Estimate, Std.Error, DF, P.Value, is_significant) |> 
  group_by(Site) |> 
  gt() |> 
  cols_hide(columns = is_significant) |> 
  tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_body(
        columns = c(Response, Predictor, Std.Estimate, Std.Error, DF, P.Value),
        rows = is_significant
      )) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_column_labels(columns = c(Response, Predictor, Std.Estimate, Std.Error, DF, P.Value))
    )

```

### SEM for species richness and evenness (clipping)

```{r}
#| label: fig-cut-final-rich-even
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 10
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and clipping on species richness and evenness, as well as their indirect effects through community biomass for the a) and c) alpine and b) and d) sub-alpine community. The numbers indicate the regression coefficients, the colour the direction (orange = positive, turquoise = negative), and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05) and linetype (solid = significant and dashed = non-significant)."

tar_load(cut_final_richness_evenness)

cut_final_richness_evenness[[1]]

```


### SEMs for natural grazing

```{r}
#| label: fig-graz-final-all-diversity
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 15
#| fig-cap: "Structural equation model showing the direct effects of the global change drivers, warming, nitrogen addition and grazing on diversity, richness, and evenness, as well as their indirect effects through community biomass for the alpine (a, c, e) and the sub-alpine community (b, d, f). The numbers indicate the regression coefficients, the colour the direction (orange = positive, turquoise = negative), and the thickness of the line relates to the strength of the effect. Significant effects are indicated by asterisks (*** for P value < 0.001, ** for 0.001 ≤ P-value < 0.01, and * for 0.01 ≤ P-value < 0.05) and linetype (solid = significant and dashed = non-significant)."

tar_load(graz_final_all_div)

graz_final_all_div[[1]]

```

### Statistics for SEMs for Shannon diversity, species richness and evenness at the alpine and sub-alpine site (grazing)

```{r}
#| label: tbl-graz-final-stats
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 10
#| tbl-cap: "SEM output for direct effects of warming, nitrogen addition, and grazing and indirect effects via community biomass on Shannon diversity, species richness and evenness. Shown are the response, predictors, standard estimate, standard error, degrees of freedome (DF) and P value. Variables in bold have a significant P value."

tar_load(graz_final_all_div)

graz_final_all_div[[2]] |> 
  mutate(diversity_index = case_when(
           str_detect(Type, "diversity") ~ "diversity",
           str_detect(Type, "richness") ~ "richness",
           str_detect(Type, "evenness") ~ "evenness"
         ),
         Site = case_when(
           str_detect(Type, "Alpine") & !str_detect(Type, "Sub-alpine") ~ "Alpine",
           str_detect(Type, "Sub-alpine") ~ "Sub-alpine"
         )) |> 
  mutate(P.Value_num = P.Value,
         is_significant = P.Value <= 0.05) |> 
  mutate(Std.Estimate = round(Std.Estimate, 2),
         Std.Error = round(Std.Error, 2),
         P.Value = if_else(P.Value_num <= 0.001, "≤ 0.001", as.character(round(P.Value_num, 3)))) |> 
  select(Site, Response, Predictor, Std.Estimate, Std.Error, DF, P.Value, is_significant) |> 
  group_by(Site) |> 
  gt() |> 
  cols_hide(columns = is_significant) |> 
  tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_body(
        columns = c(Response, Predictor, Std.Estimate, Std.Error, DF, P.Value),
        rows = is_significant
      )) |> 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_column_labels(columns = c(Response, Predictor, Std.Estimate, Std.Error, DF, P.Value))
    )

```

## Environmental affinities

### Environmentalf affinity distributions

To understand how traits respond to the different global change drivers, we analyzed the effects of warming, nitrogen addition, clipping, and community biomass on plant height and Ellenberg indicator values (@tbl-trait-stats).

```{r}
#| label: fig-affinity-distributions
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Trait distributions for plots with different clipping treatments at the alpine (upper row) and sub-alpine site (bottom row). Shown are the trait distributions for plant height and Ellenberg light, temperature, nutrient, moisture, salinity and reaction indicator values, as well as a mowing and grazing disturbance indicator. The colours indicate different levels of clipping."

# Load the individual plots
tar_load(traits_clipping_plot)

traits_clipping_plot

```

### Environmental affinity distributions statistics

```{r}
#| label: tbl-trait-stats
#| echo: false
#| warning: false
#| tbl-cap: "Statistical results for trait responses to global change drivers. Shown are the effects of warming, nitrogen addition, clipping, and community biomass on plant height and Ellenberg indicator values (light, temperature, nutrients, moisture, reaction, mowing, and grazing disturbance) at the alpine and sub-alpine sites. Significant effects (P ≤ 0.05) are shown in bold."

tar_load(trait_stats_table)

trait_stats_table

```
